### 2017-09-29 documentation/_region_.tex
adding documentation, partially implementing contributions/weights

\message{ !name(paper.tex)}\documentclass{article}
\usepackage{amsmath}
\usepackage{listings}
\usepackage{color}
\usepackage[margin=1.15in]{geometry}

\lstset{frame=tb,
  basicstyle=\footnotesize,
  language=C,
  aboveskip=2mm,
  belowskip=2mm,
  frame=none,
  keepspaces=true,
  showstringspaces=false,
  columns=flexible,
  basicstyle={\small\ttfamily},
  breaklines=false,
  breakatwhitespace=false,
  tabsize=2
}

\lstset{language=C}

\title{DTOADQ}

\begin{document}

\message{ !name(paper.tex) !offset(-3) }

  \maketitle
  \section{Introduction}
    I apologize for this being very lengthy and at times mathematically heavy,
    but there is no way of avoiding it.

  \section{SDFs}
  \begin{align*}
    S(P, R) &= P - R^2   & \text{SDF of a sphere Radius  }\\
    S(P, N) &= P \cdot N  & \text{SDF of a plane at Normal}\\
    S(P, H, R) &= <|P_{x, z}| - R, \: P_{y}> - H & \text{SDF of a torus with
                                                  Radius and Hole}\\
  \end{align*}
  \section{OpenCL}
  \section{Light Transport Introduction}
  \section{Bidirectional Path Tracing}
    All the materials in here, about path tracing, have come from Veach's PhD
    thesis. So, in a way, DTOADQ, at this stage, is just an implementation of
    BDPT as described in veach's thesis to the GPU, along with pulling from
    other resources such as Physically Based Render [PBR] 3rd ed. In order
    to understand the mathematics for BDPT, it must be worked up starting from
    light transport. So this section is about transforming the algorithms and
    mathematical models into something easily and efficiently implementable for
    the GPU, even if some of it is a rewrite of the mathematical model so that a
    programmer could understand it. In order to compute a normal of an SDF, you
    must perform a gradient approximation: the central distance onf the SDF at
    P. For example, the difference quotient:
      \[\frac{(f(x+h) - f(x))}{h}\]
    The most obvious way to calculate this is the six-point gradient; where for
    each axis you compute $f(x+h) - f(x-h)$, and normalize the result. This
    returns a vector pointing in the direction where the map's SDF changes the
    most: the normal. A four point gradient is possible; as long as each each
    point of the gradient is multiplied by $\epsilon$ before normalizing. The
    speedup from six to four is worthwhile, as computing each gradient is
    expensive as it involves a sdf map call. There are more utilities with
    gradients not yet currently provided in DTOADQ but are worth mentioning, a
    few are: an additional normal call using an alternative (higher-quality) SDF
    map can be used to compute a 'normal' bumpmap, and analyzing the gradient by
    hand allows a cheap normal computation of primitives such as a spheres.

    Rendering equation with solid angle:
  \begin{align}
    L_o(P, \omega_o) = L_o(P, \omega_o) + \int_{\Omega}f_s(P, \omega_i,
    \omega_o) L_i(P, \omega_i) cos\Theta_i d\omega_i
  \end{align}
    To get from solid angle to path, consider:
  \begin{align}
    f_s(v \rightarrow \dot{v} \rightarrow \ddot{v}) &= f_s(P, \omega_i,
                                                      \omega_o)\\
    L_e(v \rightarrow \dot{v}) &= L_e(P, \omega_o)\\
    L_e(v \rightarrow \dot{v}) &= L_e^0(v) L_e^1(v \rightarrow \dot{v})
  \end{align}
    That is, instead of looking at a point $P$ with an incoming and outgoing
    angle $(\omega_i,\:\omega_o)$, look at the path from $P_{-1}$ ($v$) to
    $P_{+1}$ ($\ddot{v}$). More accurately, it's the transport between
    three different surfaces of the scene. In terms of SDFs, the surface
    is every point on an SDF map where $|map| \leq \epsilon$. There's no longer
    any integration over a hemisphere with solid angle $\omega_i$, it's now over
    every single surface in the scene. This will be explained in detail later.
\\
    In the case of $L_e$,
    DTOADQ [at least, for now], only has support for area-light. Thus consider:

  \begin{align}
    L_e(P, \omega_o) &\equiv L_e(P) &\text{[ for area lights ]}\\
    L_e(v \rightarrow \dot{v}) &= L_e^0(v)&
  \end{align}

    The cos term, used for the projected solid angle in irradiance, no longer
    applies when dealing with area. This and $d\omega_i$ now become a
    generalized geometry term, G, that converts a PDF with respect to solid area
    using Jacobian mapping, which involves the inverse squared distance, and the
    cosine angle between the geometric normals at $v$ and $\dot{v}$:

  \begin{align}
    cos\theta_n = v \cdot N(v) \\
    G(v \leftrightarrow \dot{v}) = V(v \rightarrow \dot{v})
        \frac{(cos\theta_n cos\dot{\theta_n})}{|| v - \dot{v}||^2}
  \end{align}

    Where $V$ is the visibility test. Thus the rendering equation with respects
    to area is:
  \begin{align}
    L(v \rightarrow \dot{v}) = L_e(v \rightarrow \dot{v}) \int_ML(v \rightarrow
    \dot{v}) G(v\leftrightarrow \dot{v}) f_s(v\rightarrow \dot{v} \rightarrow
    \ddot{v}) d_A
  \end{align}
    where $A$ is the area on $M$, and $M$ is the union of all scene surfaces.
    This is known as the three-point form or the light transport equation. By
    recursively expanding$^{[1]}$, so that the three point form is integrated
    over every possible set of possible paths, this can be rewritten as the more
    usable format:
  \begin{align} I_j = \int_\Omega f_j(\bar{v}) d\mu(\bar{v}) \end{align}
    Where $\bar{v}$ is a path $z_0\rightarrow z_s \rightarrow y_t \rightarrow
    y_0 (eye \rightarrow light)$, $\Omega$ is the combination of all possible
    paths of any length, and $\mu$ is the area-product measure [the product of
    all the expanded $d_A$]. Special care has to be taken with the definition of
    path itself, as in the source code, there never exists a path $\bar{v}$,
    only eye-path $\bar{z}$ and light-path . Right now, this is analytically
    unsolvable. To compute this, we need to apply monte carlo sampling and limit
    the maximum path length:
  \begin{align}
    I_j \approx \frac{1.0f}{N}\Sigma_0^NF\\
    F = \Sigma_s\Sigma_t\frac{f_j(\bar{v}_{s, t})}{p(\bar{v}_{s, t})}
  \end{align}

    where $s$ is the light-path length, and $t$ is the eye-path length. In
    DTOADQ, they are limited to $s+t \leq 8$[TODO]. Finally, to transform
    $\bar{v}$ into two seperate paths, $\frac{f_j(\bar{v}_{s,
        t})}{p(\bar{v}_{s, t})}$ is split into two with a connection edge $c$
  \begin{align}
    \frac{f_j(\bar{v}_{s, t})}{p(\bar{v}_{s, t})} \equiv
    \frac{f^L(\bar{y})}{p(\bar{y})} c(\bar{y}_s \leftrightarrow \bar{z}_t)
    \frac{f^E(\bar{z})}{p(\bar{z})}  
  \end{align}

    And now
  \begin{align}
    F = \Sigma_s\Sigma_t \alpha^L_s c_{s, t} \alpha^E_t
  \end{align}
  \begin{align}
    \alpha^{L|E}_i =
    \begin{cases}
      \frac{L_e^0(y_0 \rightarrow y_1)}{p_A(y_1)} | \frac{W^0_e(z_0)}{p_A(z_0)},
      &i = 1,\\
      (\frac{f_s(y_{i+1} \rightarrow y_i \rightarrow y_{i-1})}{p_{\sigma}(y_i
      \rightarrow y_{i-1})} | \frac{f_s(z_{i-1} \rightarrow z_{i-2} \rightarrow
      z_{i-3}}{p_\sigma(z_{i-2} \rightarrow z_{i-1})}) \alpha^{L|E}_{i-1},
      &i > 1
    \end{cases}
  \end{align}
    There is no case for $i = 0$ as either, the path doesn't exist, or if one
    does exist, $y_0$ and $z_0$ do not contribute to the image. Veach (and many other
    sources), leave this in, but it's unnecessary and can be confusing. However,
    there may be cases where $s = 0$ ($t = 0$ is not possible without a
    physical camera lens), which is equivalent to just forward path tracing.
    Well, in either case where both paths exist or only the camera path exists,
    the connection strategy below describes how to connect the edges of the two
    paths
  \begin{align}
    c_{s, t} =
    \begin{cases}
      L_e(z_{-2} \rightarrow z_{-1}), &s = 0, t > 0,\\
      f_s(y_{-2} \rightarrow y_{-1} \rightarrow z_{-1}) G(y_{-1} \leftrightarrow
      z_{-1}) f_s(y_{-1} \rightarrow z_{-1} \rightarrow z_{-2}), &s = s, t > 0
    \end{cases}
  \end{align}
    One immediately obvious optimization that could be made with this model is
    in regards to the visibility check in the geometric term, if you were to
    expand $I_j$, and take the V term from outside the G function
    [$G(v \rightarrow
    \dot{v}) V(v \rightarrow \dot{v})$],
    it would be made obvious that all the visibility checks can be
    cancelled out for the entire equation, except for the connection term $c$.
    The expansion is not shown as it is very lengthy, and unnecessary. Paths
    $\bar{y}_{0} \ldots \bar{y}_s$ and $\bar{z}_0 \ldots \bar{z}_t$ being
    visible is intuitively obvious as we are already generating these paths.
    Specifically, the definition of $V$ is
  \begin{align}
    V(v \rightarrow \dot{v}) = ||v - \dot{v}|| \leq
    m(v,\:\overrightarrow{v\:\dot{v}})
  \end{align}
    where $m$ is a march through the SDF map. Since the paths are being
    generated using march. A visibility check is only necessary for the
    connection edge
    \begin{align*}
      ||v - \dot{v}|| = m(v, \overrightarrow{v, \dot{v}})\\
      \:\:\:\text{for all paths but } y_{-1} \rightarrow z_{-1} 
    \end{align*}
\\
    Moving on, for $\alpha^{L|E}_i$, these are equivalent for both $L$ and $E$,
    the problem is that the $\bar{y}$ path is generated in perspective of the
    light source, so we have to call the BSDF in perspective of the camera; the
    path is flipped. Handling $\bar{y}$ is significantly easier than $\bar{z}$,
    as $\bar{y}$ is precomputed. For $\bar{z}$, the next vertex $z_{+1}$
    ($\omega_o$ in the perspective of SA), is unknown. For example $\alpha^L_1$
    in OpenCL looks like:

    \begin{lstlisting}
      Vertex* V0 = light_path.vertices - 1, * V1 = light_path.vertices;
      float sigma_pdf = Light_PDF(V1.origin, &light);
      float g = Geometric_Term(V0.origin, V1.origin, V0.normal, V1.normal);
      float light_pdf = sigma_pdf * g;
      float Le = light.colour;
      light_contrib[1] = Le / light_pdf;
    \end{lstlisting}

    For $c_{s, t}$, there is no cases for $s = s, t = 0$ as
    the lens does not have a physical existence in the SDF map. $p_A$ is
    the PDF w.r.t. area, for the light path; we only need to take the PDF of the
    solid angle of a sphere
  \begin{align}
    p_A^L(y_0 \rightarrow y_1) = p_{\sigma}^L(y_0 \rightarrow y_1) G(v
    \leftrightarrow \dot{v})\\
    p_{\sigma}^L(y_0 \rightarrow y_1) = \frac{1.0f}{\tau * (1.0f -
    \sqrt{\frac{R^2}{|y_0 - y_1|^2}})}
  \end{align}
    where $R$ is the radius of the sphere, note $y_0$ is a point on the surface
    of the sphere, and not its origin.
    \\$p_A^E(x_0 \rightarrow x_1)$ doesn't exist, as there is no physical lens
    on the scene. [TODO - this might cahnge later]. One of the benefits of this,
    is that there are no special cases for pathes where $t = 0$, which is
    impossible to handle on the GPU as the path could contribute to any
    arbitrary pixel [directly hitting the lens].
  \begin{align}
    P_A^E = 1.0f
  \end{align}
    The $L^0_e$ is contribution from the light source. As of right now, DTOADQ
    only supports area lights, so $L^0_e$ is set a constant for the emitter's
    spectrum value.
    The $W_e^0$ is contribution from the eye source, that is, the lens. Since
    the lens does not exist in the scene, as of now [TODO - this might change
    later], there is no $W_e^0$. While this should be implemented in the future,
    a hidden benefit of the current model is that there are no special cases for
    paths where $t = 0$, which is impossible to handle on the GPU as any such
    path could contribute to any arbitrary pixel [the light path is, after all,
    directly hitting the lens]. So this has now become:
  \begin{align}
    \alpha_i^E = \Sigma_{i\leq 0} 
  \end{align}
  The amount that each path contributes to F differ greatly, so naively
  weighing each contribution uniformly by something like 1.0f, or even, by
  inverse path length, will producy very noisy images, and would not be
  worth the additional computation time that BDPT requires. In order to get good
  samples from the contributions, they must be combined in an optimal manner
  using Multiple Importance Sampling: each contribution is weighted by the PDF
  of the entire path itself:
  \begin{align*}
    F &\equiv \Sigma_s\Sigma_tW_{s, t}C_{s, t}\\
    C_{s, t} &= \alpha^L_s c_{s, t} \alpha^E_t
  \end{align*}
    where $W_{s, t}$ is the MIS weight of path $\bar{y} \leftarrow \bar{z}$, and
    $C_{s, t}$ is the unweighted contribution of the same path. For clarity,
    $C^U_{s, t}$ as described below is the unweighted contribution of the
    algorithm:
  \begin{align}
    C^u_{s, t} = (\alpha^L_s c_{s, t} \alpha^E_t) \equiv
    \frac{f_j(\bar{v})}{p(\bar{v})}
  \end{align}
  thus
  \begin{align}
    F = \Sigma_s\Sigma_tW_{s, t}C^u_{s, t}
  \end{align}
    The formula Veach gives for Multiple Importance Sampling on BDPT is as so:
  \begin{align}
    MIS_{st}(\bar{x}) = \frac{P_s(\bar{x})}{\Sigma_i{P_i(\bar{x})}} 
  \end{align}
    Conceptually, the numerator $P_s(\bar{x})$ is the path density that was
    generated, while the denominator is the path density of other connection
    strategies that could could have, in theory,$^{[1]}$ created the path. The
    straight forward implementation of this has a lot of problems gone
    unresolved in Veach's thesis; PBR 3rd Ed, chapter 16 pg 1014-1016 addresses
    and describe these problems. Primarily, $P_i(\bar{x})$ will overflow
    CLFloat (due to the distance in the Geometric term), and the straight
    forward implementation is very long, hard to debug, and has a poor time
    complexity. But the important part is the end result of their solution:

  \begin{align}
    r_i (\bar{x}) = 
    \begin{cases}
      1.0f, & i = s,\\
      \frac{\overleftarrow{P}(\bar{x_i})}{\overrightarrow{P}(\bar{x_i})} *
      r_{i+1}(\bar{x}), & i < s,\\
      \frac{\overrightarrow{P}(\bar{x_{i-1}})}{\overleftarrow{P}(\bar{x_{i-1}})}
      * r_{i-1}(\bar{x}), & i > s
    \end{cases}
  \end{align}

    However, DTOADQ can simplify a bit further; unlike above, there are two
    paths for which to evaluate. Specifically, in the above term, for where
    $i < s$, we have
    $\frac{\overleftarrow{p}(\bar{x}_i)}{\overrightarrow{p}(\bar{x}_i)}$, 
    which is the light path (i is iterating towards light-length s). The eye
    contribution is inversed from this, (also iterating towards s, hence why
    $\bar{x}_{i-1}$ is used). Another way of saying it, is $\bar{x} = \bar{y} ||
    \bar{z}_{rev}$.
    In DTOADQ's case, the weights are calculated in
    the same direction relative to its originator; in this case the perspective
    of the eye path.
  \begin{align}
    \mathcal{W}_i(\bar{x}) \equiv r_i(\bar{x})\\
    \mathcal{W}_i (\bar{x}) = 
    \begin{cases}
      1.0f, & i = 0,\\
      \frac{\overleftarrow{p_i}(\bar{x_i})}{\overrightarrow{p_i}(\bar{x_i})} *
      \mathcal{W}_{i-1}(\bar{x}), & i > 0
    \end{cases}
  \end{align}
    And then the expanded form of $\mathcal{W}_i(\bar{x})$ looks like

  \begin{align}
    \mathcal{W}_i(\bar{x}) = \Pi_{n=2}^i(\frac{\overleftarrow
    {p_{\sigma}}(x_n)\overleftarrow {G}(x_n)}
    {\overrightarrow{p_{\sigma}}(x_n)\overrightarrow{G}(x_n)})
    * \frac{\overleftarrow{p_A}(x_1)}{\overrightarrow{p_A}(x_1)}
  \end{align}

    And now the MIS looks like:
  \begin{align}
    MIS_{s, t}(\bar{y}, \bar{z}) =
    \frac{1.0f}{\mathcal{W}_s(\bar{y}) + \mathcal{W}_t(\bar{z}) + 1.0f}
  \end{align}
    Thus, the final rendering equation for DTOADQ looks like such:

  \begin{align}
    P = \frac{1.0f}{N} \Sigma_{0}^{\infty} \Sigma_{s \le 0} \Sigma_{t \le 0}
    \alpha^L_s c_{s,t} \alpha^E_t MIS_{s, t}(\bar{y}, \bar{z})
  \end{align}

  \begin{align}
    P_i(\bar{x}_{s, t}) = P^L_s p^E_t
  \end{align}

  \begin{align}
    p^{L|E}_i(\bar{x}) =
    \begin{cases}
      1.0f, & i = 0,\\
      P_A(x_i), & i = 1,\\
      P_{\sigma}(x_{i-1} \rightarrow x_i) G(x_{i-1} \leftrightarrow x_i)
      P^{L|E}_{i-1}, & i > 1
    \end{cases}
  \end{align}

  \section{Video/Image Emitter}

  \section{Bibliography}
  [1]
  \begin{align*}
    \Sigma_{k=1}^{\infty} \int_{M^{k+1}} L_e(v_0 \rightarrow v_1) G(v_0
    \leftrightarrow v_1)
    \Pi_{i=1}^{k-1} F_s(v_{i-1} \rightarrow v_i \rightarrow v_{i+1}) G(v_i
    \leftrightarrow v_{i+1})\\
    W_e^j(v_{k-1} \rightarrow v_k) dA(v_0) \ldots dA(v_k)
  \end{align*}
\end{document}

\message{ !name(paper.tex) !offset(-368) }

### 2017-09-29 documentation/paper.tex
adding documentation, partially implementing contributions/weights

--- 2017-09-29 documentation/_region_.tex
@@
-\message{ !name(paper.tex)}\documentclass{article}
+\documentclass{article}
 \usepackage{amsmath}
@@
 \begin{document}
-
-\message{ !name(paper.tex) !offset(-3) }
-
   \maketitle
@@
     does exist, $y_0$ and $z_0$ do not contribute to the image. Veach (and many other
-    sources), leave this in, but it's unnecessary and can be confusing. However,
-    there may be cases where $s = 0$ ($t = 0$ is not possible without a
+    sources), leave this in, but it's unnecessarily confusing. A special case
+    needs to be handled, unfortunately, for $\alpha^E_0$ as it is being
+    generated on the fly while $\bar{z}$ is being constructed. The two options
+    is to unroll the first iteration of the construction loop, or just allow the
+    special case to exist. The latter is the better choice, as all kernels will
+    enter and exit the special case at the same time, and the GPU might just
+    unroll the loop anyways.
+    There may be cases where $s = 0$ ($t = 0$ is not possible without a
     physical camera lens), which is equivalent to just forward path tracing.
@@
     $\bar{y}_{0} \ldots \bar{y}_s$ and $\bar{z}_0 \ldots \bar{z}_t$ being
-    visible is intuitively obvious as we are already generating these paths.
-    Specifically, the definition of $V$ is
+    visible is intuitively obvious as the paths have been generated using the
+    same technique used to check for visibility; raymarching. Specifically, the
+    definition of $V$ is
   \begin{align}
@@
   \end{align}
-    where $m$ is a march through the SDF map. Since the paths are being
-    generated using march. A visibility check is only necessary for the
-    connection edge
+    where $m$ is a march through the SDF map. A visibility check is only
+    necessary for the connection edge
     \begin{align*}
@@
 \end{document}
-
-\message{ !name(paper.tex) !offset(-368) }

### 2017-09-29 dtoadqkernel.c

@@ -85,13 +85,8 @@ typedef struct T_Vertex {
   int mat_index;
 } Vertex;
 
-typedef struct T_Edge {
-  float3 wi, wo;
-} Edge;
-
 typedef struct T_Subpath {
   Vertex vertices[MAX_DEPTH];
-  Edge   edges   [MAX_DEPTH];
   uint length;
 } Subpath;
 
@@ -216,22 +211,6 @@ SampledPt March ( int avoid, Ray ray, SCENE_T(si, Tx)) {
 // -----------------------------------------------------------------------------
 // --------------- GRAPHIC FUNCS -----------------------------------------------
 float3 Normal ( float3 p, SCENE_T(s, t)) {
-  /* In order to do a normal, using SDFs, you must perform a gradient
-       approximation: the central difference on the SDF at P, for example the
-        difference quotient: (f(x+h) - f(x))/h
-     The most obvious way to calculate this is the six-point gradient; where
-       for each axis you compute f(x+h) - f(x-h), and normalize the result,
-       this returns a vector pointing in the direction where the SDF map
-       changes the most, so the gradient is the same as a normal.
-     A 4 point gradient is possible, you just have to multiply each point
-       of the gradient by epsilon before normalizing. Computing these gradients
-       is expensive, because you have to map multiple times, so the speedup
-       is well worth the (rather unnoticeable) accuracy.
-     It's worth noting that with an additional normal, you can also compute a
-       bumpmap, and that with simple object you can calculate the gradient
-       by hand rather than this general-purpoe solution. But that's not an
-       option with dtoadq.
-  */
   float2 e = (float2)(1.0f, -1.0f)*0.5773f*0.0005f;
   return normalize(
     e.xyy*Map(-1, p + e.xyy, s, t).dist + e.yyx*Map(-1, p + e.yyx, s, t).dist +
[…]

### 2017-10-09 documentation/_region_.tex
documentation & path tracing

@@
 \usepackage{listings}
+\usepackage{amssymb}
 \usepackage{color}
+\usepackage{graphicx}
+\graphicspath{ {../screenshots/} }
 \usepackage[margin=1.15in]{geometry}
@@
   \section{Introduction}
-    I apologize for this being very lengthy and at times mathematically heavy,
-    but there is no way of avoiding it.
 
@@
       \[\frac{(f(x+h) - f(x))}{h}\]
+
+    Calculating probability distribution functions is a very vital component in
+    Monte Carlo Sampling. The PDF states the probability of a random variable to
+    take the form of a given input. Its usefulness comes when contributions from
+    many Monte Carlo samples have a different weight. That is, one sample may
+    have a higher probability distribution, and thus a higher contribution to
+    the path. That's the strategy behind Monte Carlo;
+    \begin{align}
+      \int f(x) \approx \frac{1}{N}\Sigma_0^N\frac{f(x)}{p(x)}
+    \end{align}
+    The PDF applies to everything from path contributions, solid angles of a
+    BRDF, uniformly sampling a sphere, etc. I had a difficult time understanding
+    these, and more importantly, how to calculate them for a given formula. The
+    biggest conceptual barrier is that, for example, given a solid angle in a
+    hemisphere $\Omega$, the probability for a random variable X to equal
+    $\omega$ is 0. But this isn't what's happening here, this is all about the
+    probability that X could potentially equal $\omega$ when compared to other values
+    in $\Omega$. It's all about comparing the probabilities of two potential
+    $\omega$s with each other.
+
+    I still have a hard time conceptualizing these things in my head, but the
+    good thing is, is that calculating these are relatively easy. There's a lot
+    that could be talked about, but the important thing is that I've found a
+    reliable three-step process. The first is to
+    find the cumulative distribution function $CDF = Pr(X \leq x)$. Where $Pr$
+    is the probability, $X$ is a uniformly random variable, and $x$ is a known
+    constant. In other words, this gives the probability that, for a formula
+    $P$, where $X \in \Omega$, and $\Omega$ is the set of all possible values
+    for $CDF$,
+    that $X \leq x$. The second step is to take the derivative of this, so for
+    PDF $p$
+    \begin{align}
+      p(x) = \frac{dCDF(x)}{dx}
+    \end{align}
+    And the final step is to verify that the PDF is correct by integrating it
+    over $\Omega$, this final value should equal 1. This makes sense intuitively
+    -- it is a probability after all, so the sum of the probabilities of every
+    possible input must be 1.
+    For an example, how I calculated the PDF of an area light source
+      $L^0_e(v) = L_o(P, \omega_o)$ 
+    is quite simple. Firstly,the area light is defined as a sphere, and thus our
+    set of all possible values is correlated to surface area:
+    $\Omega = [0,\:2.0f*\tau*R^2]$.
+    Thus for a given constant $x$, and a uniformly random $X$,
+    \begin{align}
+      CDF(x) = Pr(X \leq x) = \frac{x}{2.0f*\tau*R^2}\\
+      p(x) = \frac{dCDF(x)}{dx} = \frac{1.0f}{2.0f*\tau*R^2}\\
+      \int_0^{\Omega}p(x)dx = 1.0f
+    \end{align}
+    In code
+    \begin{lstlisting}
+      float Light_PDF ( float3 P, Emitter* m ) {
+        float rad_sqr = m->radius*m->radius;
+        return 1.0f/(2.0f*TAU*rad_sqr);
+      }
+    \end{lstlisting}
+    The more interesting aspect, in this example, is actually generating the
+    uniformly distributed samples of a sphere. The naive method gives incorrect
+    distributions.TODO {page 774}
+
     The most obvious way to calculate this is the six-point gradient; where for
@@
     does exist, $y_0$ and $z_0$ do not contribute to the image. Veach (and many other
-    sources), leave this in, but it's unnecessary and can be confusing. However,
-    there may be cases where $s = 0$ ($t = 0$ is not possible without a
+    sources), leave this in, but it's unnecessarily confusing. A special case
+    needs to be handled, unfortunately, for $\alpha^E_0$ as it is being
+    generated on the fly while $\bar{z}$ is being constructed. The two options
+    is to unroll the first iteration of the construction loop, or just allow the
+    special case to exist. The latter is the better choice, as all kernels will
+    enter and exit the special case at the same time, and the GPU might just
+    unroll the loop anyways.
+    There may be cases where $s = 0$ ($t = 0$ is not possible without a
     physical camera lens), which is equivalent to just forward path tracing.
+    There is also to consider, what happens on $s = 1$ and $t = 1$, %TODO
     Well, in either case where both paths exist or only the camera path exists,
@@
       f_s(y_{-2} \rightarrow y_{-1} \rightarrow z_{-1}) G(y_{-1} \leftrightarrow
-      z_{-1}) f_s(y_{-1} \rightarrow z_{-1} \rightarrow z_{-2}), &s = s, t > 0
+      z_{-1}) f_s(y_{-1} \rightarrow z_{-1} \rightarrow z_{-2}), &s > 0, t > 0
     \end{cases}
   \end{align}
+    One question to be made is if the strategy $s = 0, t > 0$ is worthwhile to
+    compute; the most noticeable effect are for caustics, light sources in
+    direct view of an emitter should be emitted from animations in this current
+    state of DTOADQ. Another idea to consider is that under real physics
+    circumstances, the photons are still allowed to bounce after hitting a
+    source of light, but the contribution from these cases are negligible, and
+    so most likely not worthwhile to implement. Anyways, the $s = 0$
+    strategy is a worthwhile  calculation under these circumstances
+
     One immediately obvious optimization that could be made with this model is
@@
   inverse path length, will producy very noisy images, and would not be
-  worth the additional computation time that BDPT requires. In order to get good
+  worth the additional computation time that BDPT requires, as shown below.
+  \begin{align}
+ \includegraphics{bdpt_unweighted}
+  \end{align}
+  In order to get good
   samples from the contributions, they must be combined in an optimal manner
@@
   \end{align}
-    The formula Veach gives for Multiple Importance Sampling on BDPT is as so:
+    The formula for Multiple Importance Sampling on BDPT is as so:
   \begin{align}
@@
 
+  \section{Future Improvements}
+
+  Every emitter is chosen uniformly, however a better system would weight the
+  probability of each emitter by
+  \[
+    p_l = \frac{\phi_lr_l}{\Sigma_i\phi_ir_l}
+  \]
+  Thus emitters that have a higher contribution to the scene, that is, a larger
+  emittance and radius, will have a higher probability to be selected.
+
   \section{Bibliography}
@@
 
-\message{ !name(paper.tex) !offset(-368) }
+\message{ !name(paper.tex) !offset(-461) }

### 2017-10-09 documentation/paper.tex
documentation & path tracing

@@
     physical camera lens), which is equivalent to just forward path tracing.
+    There is also to consider, what happens on $s = 1$ and $t = 1$, %TODO
     Well, in either case where both paths exist or only the camera path exists,

### 2017-10-09 documentation/paper/images.pl

@@ -1,0 +1,858 @@
+# LaTeX2HTML 2017.2 (Released Jan 23, 2017)
+# Associate images original text with physical files.
+
+
+$key = q/(omega_i,:omega_o);MSF=1.6;AAT/;
+$cached_env_img{$key} = q|<IMG
+ WIDTH="59" HEIGHT="32" ALIGN="MIDDLE" BORDER="0"
+ SRC="|."$dir".q|img35.png"
+ ALT="$ (\omega_i,\:\omega_o)$">|; 
+
+$key = q/A;MSF=1.6;AAT/;
+$cached_env_img{$key} = q|<IMG
+ WIDTH="16" HEIGHT="14" ALIGN="BOTTOM" BORDER="0"
+ SRC="|."$dir".q|img52.png"
+ ALT="$ A$">|; 
+
+$key = q/CDF;MSF=1.6;AAT/;
+$cached_env_img{$key} = q|<IMG
+ WIDTH="43" HEIGHT="15" ALIGN="BOTTOM" BORDER="0"
+ SRC="|."$dir".q|img17.png"
+ ALT="$ CDF$">|; 
+
+$key = q/CDF=Pr(Xleqx);MSF=1.6;AAT/;
+$cached_env_img{$key} = q|<IMG
+ WIDTH="141" HEIGHT="32" ALIGN="MIDDLE" BORDER="0"
+ SRC="|."$dir".q|img11.png"
+ ALT="$ CDF = Pr(X \leq x)$">|; 
+
+$key = q/C^U_{s,t};MSF=1.6;AAT/;
+$cached_env_img{$key} = q|<IMG
+ WIDTH="31" HEIGHT="35" ALIGN="MIDDLE" BORDER="0"
+ SRC="|."$dir".q|img117.png"
+ ALT="$ C^U_{s, t}$">|; 
+
+$key = q/C_{s,t};MSF=1.6;AAT/;
+$cached_env_img{$key} = q|<IMG
+ WIDTH="31" HEIGHT="30" ALIGN="MIDDLE" BORDER="0"
+ SRC="|."$dir".q|img116.png"
+ ALT="$ C_{s, t}$">|; 
+
+$key = q/E;MSF=1.6;AAT/;
+$cached_env_img{$key} = q|<IMG
+ WIDTH="17" HEIGHT="14" ALIGN="BOTTOM" BORDER="0"
[…]

### 2017-10-09 documentation/paper/labels.pl

@@ -1,0 +1,13 @@
+# LaTeX2HTML 2017.2 (Released Jan 23, 2017)
+# Associate labels original text with physical files.
+
+
+1;
+
+
+# LaTeX2HTML 2017.2 (Released Jan 23, 2017)
+# labels from external_latex_labels array.
+
+
+1;
+

### 2017-10-09 dtoadqkernel.c

@@ -1,11 +1,13 @@
-#define MAX_DEPTH 8
+#define MAX_DEPTH 5
 #define MARCH_DIST //%MARCH_DIST.0f
 #define MARCH_REPS //%MARCH_REPS
+#define MAT_LEN 4
 
 #define TEXTURE_T __read_only image2d_array_t
 #define SCENE_T(S, T) SceneInfo* S, TEXTURE_T T
 // EMIT_MAT = -120 - light_index
 #define EMIT_MAT -120
+#define REMIT(I) REmission(EMIT_MAT - I, si->debug_values, si->time)
 
 #define DFLT2(V) ((V).x), ((V).y)
 #define DFLT3(V) ((V).x), ((V).y), ((V).z)
@@ -13,6 +15,7 @@
 
 #define writeln(X)     if (Is_Debug()){printf(X "\n");                         }
 #define writeint(X)    if (Is_Debug()){printf(#X " %d\n",                 (X));}
+#define writeptr(X)    if (Is_Debug()){printf(#X " %p\n",                 (X));}
 #define writefloat(X)  if (Is_Debug()){printf(#X " %f\n",                 (X));}
 #define writefloat2(X) if (Is_Debug()){printf(#X " %f, %f\n",        DFLT2(X));}
 #define writefloat3(X) if (Is_Debug()){printf(#X " %f, %f, %f\n",    DFLT3(X));}
@@ -82,7 +85,7 @@ SampledPt SampledPt_From_Origin(float3 origin) {
 
 typedef struct T_Vertex {
   float3 colour, origin, normal;
-  int mat_index;
+  Material* material;
 } Vertex;
 
 typedef struct T_Subpath {
@@ -93,11 +96,11 @@ typedef struct T_Subpath {
 typedef struct T_SceneInfo {
   float time;
   // __read_only image2d_array_t textures; IMAGES CANT BE USED AS FIELD TYPES:-(
-  __global Material* materials;
+  Material* materials;
   float3 debug_values;
   float rng;
 } SceneInfo;
-SceneInfo New_SceneInfo(float time, __global Material* materials,
+SceneInfo New_SceneInfo(float time, Material* materials,
                         float3 debug_values, float rng){
   SceneInfo si;
   si.time         = time;
@@ -264,7 +267,6 @@ float3 Hemisphere_Direction ( SceneInfo* si ) {
   return To_Cartesian(sin_theta, cos_theta, phi);
 }
 
-
 //// from PBRT
[…]

### 2017-10-09 raytracekernel.c

@@ -73,6 +73,13 @@ typedef struct T_SampledPt {
   int material_index;
 } SampledPt;
 // -----------------------------------------------------------------------------
+// --------------- GENERAL UTILITIES      --------------------------------------
+float Distance(float3 u, float3 v) {
+  float x = u.x-v.x, y = u.y-v.y, z = u.z-v.z;
+  return sqrt(x*x + y*y + z*z);
+}
+float sqr(float t) { return t*t; }
+// -----------------------------------------------------------------------------
 // --------------- MAP GEOMETRY FUNCTIONS --------------------------------------
 
 //---MAP GEOMETRY INSERTION POINT---
@@ -115,7 +122,7 @@ SampledPt Map ( int a, float3 origin, SCENE_T(si, Tx) ) {
 
 // -----------------------------------------------------------------------------
 // --------------- GRAPHIC FUNCS -----------------------------------------------
-float3 Normal ( float3 p, float tm, SCENE_T(si, Tx) ) {
+float3 Normal ( float3 p, SCENE_T(si, Tx) ) {
   float2 e = (float2)(1.0f, -1.0f)*0.5773f*0.0005f;
   return normalize(
     e.xyy*Map(-1, p + e.xyy, si, Tx).dist +
@@ -155,14 +162,85 @@ SampledPt March ( int avoid, Ray ray, SCENE_T(si, Tx) ) {
 //----POSTPROCESS---
 //%POSTPROCESS
 
-SampledPt Colour_Pixel(Ray ray, SCENE_T(si, Tx) ){
+float Visibility_Ray(float3 orig, float3 other, float sphere_radius,
+                     SCENE_T(si, Tx)) {
+  float theoretical = Distance(orig, other) - sphere_radius;
+  float3 dir = normalize(other - orig);
+  orig += dir*(MARCH_ACC*2.0f + 0.2f);
+  SampledPt ptinfo = March(-1, (Ray){orig, dir}, si, Tx);
+  float actual = ptinfo.dist + MARCH_ACC + 0.4f;
+  return 1.0f*(actual >= theoretical);
+}
+
+float3 BRDF_Specular(float3 L, float3 V, float3 N, float3 H, float3 col) {
+  float roughness = 0.4f, metallic = 0.3f, fresnel = 1.0f;
+  float3 F, G, D;
[…]

### 2017-10-18 dtoadqkernel.c
almost done with bpt just need mis

@@ -15,6 +15,7 @@
 
 #define writeln(X)     if (Is_Debug()){printf(X "\n");                         }
 #define writeint(X)    if (Is_Debug()){printf(#X " %d\n",                 (X));}
+#define writeint2(X)   if (Is_Debug()){printf(#X " %d, %d\n",             (X));}
 #define writeptr(X)    if (Is_Debug()){printf(#X " %p\n",                 (X));}
 #define writefloat(X)  if (Is_Debug()){printf(#X " %f\n",                 (X));}
 #define writefloat2(X) if (Is_Debug()){printf(#X " %f, %f\n",        DFLT2(X));}
@@ -59,6 +60,7 @@ typedef struct T_SharedInfo {
   unsigned char clear_img;
   unsigned long finished_samples;
   unsigned char spp;
+  uint2 rng_state;
 } SharedInfo;
 
 __constant float PI   = 3.141592653589793f;
@@ -98,28 +100,41 @@ typedef struct T_SceneInfo {
   // __read_only image2d_array_t textures; IMAGES CANT BE USED AS FIELD TYPES:-(
   Material* materials;
   float3 debug_values;
-  float rng;
+  uint2 rng_state;
 } SceneInfo;
 SceneInfo New_SceneInfo(float time, Material* materials,
-                        float3 debug_values, float rng){
+                        float3 debug_values, uint2 rng_state) {
   SceneInfo si;
   si.time         = time;
   si.materials    = materials;
   si.debug_values = debug_values;
-  si.rng          = rng;
+  si.rng_state    = rng_state;
   return si;
 }
 // -----------------------------------------------------------------------------
-// --------------- GENERAL FUNCTIONS      --------------------------------------
+// --------------- RANDOM FUNCTIONS       --------------------------------------
+/*
+  Using a high quality uniform RNG is very important for monte carlo, thus I
+    use the MWC64X.
+    http://cas.ee.ic.ac.uk/people/dt10/research/rngs-gpu-mwc64x.html
+
+  The Warp Geneartor is, while superior, does not allow divergence, which makes
+  it practically useless. For example, can't really know how much random numbers
+  you'll need for a path you haven't generated yet of unknown length.
+*/
+
[…]

### 2017-10-18 raycastkernel.c

@@ -5,25 +5,7 @@
 #define TEXTURE_T __read_only image2d_array_t
 #define SCENE_T(S, T) SceneInfo* S , TEXTURE_T T
 
-#define DFLT2(V) ((V).x), ((V).y)
-#define DFLT3(V) ((V).x), ((V).y), ((V).z)
-#define DFLT4(V) ((V).x), ((V).y), ((V).z), ((V).w)
-
-#define writeln(X)     if (Is_Debug()){printf(X "\n");                         }
-#define writeint(X)    if (Is_Debug()){printf(#X " %d\n",                 (X));}
-#define writefloat(X)  if (Is_Debug()){printf(#X " %f\n",                 (X));}
-#define writefloat2(X) if (Is_Debug()){printf(#X " %f, %f\n",        DFLT2(X));}
-#define writefloat3(X) if (Is_Debug()){printf(#X " %f, %f, %f\n",    DFLT3(X));}
-#define writefloat4(X) if (Is_Debug()){printf(#X " %f, %f, %f, %f\n",DFLT4(X));}
-
 __constant float MARCH_ACC = //%MARCH_ACC.0f/1000.0f;
-// -----------------------------------------------------------------------------
-// --------------- DEBUG -------------------------------------------------------
-// Variadic functions not supported, so this is best you get :-(
-bool Is_Debug ( ) {
-  return get_global_id(0) == get_global_size(0)/2 &&
-         get_global_id(1) == get_global_size(1)/2;
-}
 // -----------------------------------------------------------------------------
 // --------------- GPU-CPU STRUCTS ---------------------------------------------
 typedef struct T_Camera {
@@ -73,13 +55,6 @@ typedef struct T_SampledPt {
   int material_index;
 } SampledPt;
 // -----------------------------------------------------------------------------
-// --------------- GENERAL UTILITIES      --------------------------------------
-float Distance(float3 u, float3 v) {
-  float x = u.x-v.x, y = u.y-v.y, z = u.z-v.z;
-  return sqrt(x*x + y*y + z*z);
-}
-float sqr(float t) { return t*t; }
-// -----------------------------------------------------------------------------
[…]

### 2017-10-22 documentation/paper.tex
almost done with bpt just need mis
working on bdpt

@@
   \begin{align}
-    P = \frac{1.0f}{N} \Sigma_{0}^{\infty} \Sigma_{s \le 0} \Sigma_{t \le 0}
+    P = \frac{1.0f}{N} \Sigma_{0}^{N} \Sigma_{s \le 0} \Sigma_{t \le 0}
     \alpha^L_s c_{s,t} \alpha^E_t MIS_{s, t}(\bar{y}, \bar{z})
@@
 
+        // What's defined: V1 -> V2 [we're evaluating V1] This is for a lens.
+        // So, it's ultimately completely useless until an
+        // actual lens exists in the SDF scene. Note that this isn't an actual
+        // point on a surface of the scene (that's V2, next up for eval), this
+        // is just an infinitesimal point based off where the camera currently
+        // is, just like a delta light. So while the point actually exists on a
+        // scene, the evaluation is on a "delta camera"
+
   \section{Bibliography}

### 2017-10-22 dtoadqkernel.c

@@ -65,6 +65,7 @@ typedef struct T_SharedInfo {
 
 __constant float PI   = 3.141592653589793f;
 __constant float IPI  = 0.318309886183791f;
+__constant float IPI2 = 0.159154943091895f;
 __constant float TAU  = 6.283185307179586f;
 __constant float ITAU = 0.159154943091895f;
 // -----------------------------------------------------------------------------
@@ -249,37 +250,51 @@ float3 To_Cartesian ( float sin_theta, float cos_theta, float phi ) {
   return (float3)(cos(phi)*sin_theta, sin(phi)*sin_theta, cos_theta);
 }
 
-float2 Sample_Concentric_Disk ( float2 u ) {
-  float2 offset = 2.0f*u - (float2)(1.0f);
-  if ( offset.x == 0.0f && offset.y == 0.0f ) return (float2)(0.0f);
+// from PBRT 3rd ed
+float2 Concentric_Sample_Disk(SceneInfo* si) {
+  float2 u = Uniform_Sample2(si);
+  // maps to [-1, 1]^2
+  float2 offset = 2.0f * u - (float2)(1.0f, 1.0f);
+  if ( offset.x == 0.0f && offset.y == 0.0f )
+    return (float2)(0.0f);
 
   float theta, r;
   if ( fabs(offset.x) > fabs(offset.y) ) {
     r = offset.x;
-    theta = (PI/4.0f)*(offset.y/offset.x);
+    theta = (PI/4.0f) * (offset.y/offset.x);
   } else {
     r = offset.y;
-    theta = (PI/2.0f) - (PI/4.0f)*(offset.x/offset.y);
+    theta = (PI/2.0f) * (offset.x/offset.y);
   }
-  return r*(float2)(cos(theta), sin(theta));
+  return r * (float2)(cos(theta), sin(theta));
 }
 
-float3 Sample_Cosine_Hemisphere ( float2 u ) {
-  float2 d = Sample_Concentric_Disk(u);
-  float z = sqrt(fmax(0.0f, 1.0f - SQR(d.x) - SQR(d.y)));
-  return (float3)(d.x, d.y, z);
+// from PBRT 3rd ed
+float3 Sample_Cosine_Hemisphere ( SceneInfo* si ) {
+  const float2 d = Concentric_Sample_Disk(si);
+  float phi = sqrt(fmax(0.0f, 1.0f - d.x*d.x - d.y*d.y));
[…]

### 2017-10-22 raycastkernel.c

@@ -5,6 +5,22 @@
 #define TEXTURE_T __read_only image2d_array_t
 #define SCENE_T(S, T) SceneInfo* S , TEXTURE_T T
 
+#define DFLT2(V) ((V).x), ((V).y)
+#define DFLT3(V) ((V).x), ((V).y), ((V).z)
+#define DFLT4(V) ((V).x), ((V).y), ((V).z), ((V).w)
+
+#define writeln(X)     if (Is_Debug()){printf(X "\n");                         }
+#define writeint(X)    if (Is_Debug()){printf(#X " %d\n",                 (X));}
+#define writefloat(X)  if (Is_Debug()){printf(#X " %f\n",                 (X));}
+#define writefloat2(X) if (Is_Debug()){printf(#X " %f, %f\n",        DFLT2(X));}
+#define writefloat3(X) if (Is_Debug()){printf(#X " %f, %f, %f\n",    DFLT3(X));}
+#define writefloat4(X) if (Is_Debug()){printf(#X " %f, %f, %f, %f\n",DFLT4(X));}
+
+bool Is_Debug ( ) {
+  return get_global_id(0) == get_global_size(0)/2 &&
+    get_global_id(1) == get_global_size(1)/2;
+}
+
 __constant float MARCH_ACC = //%MARCH_ACC.0f/1000.0f;
 // -----------------------------------------------------------------------------
 // --------------- GPU-CPU STRUCTS ---------------------------------------------
@@ -89,7 +105,7 @@ SampledPt Map ( int a, float3 origin, SCENE_T(si, Tx) ) {
   for ( int i = 0; i != EMITTER_AMT; ++ i ) {
     Emitter e = REmission(i, si->debug_values, si->time);
     float dist = sdSphere(origin - e.origin, e.radius);
-    MapUnionG(a, &res, dist, 0, (float3)(1.0f, 0.9f, 0.8f)*e.emission);
+    MapUnionG(a, &res, dist, -100 - i, (float3)(1.0f, 0.9f, 0.8f)*e.emission);
   }
 
   return res;
@@ -102,15 +118,30 @@ SampledPt March ( int avoid, Ray ray, SCENE_T(si, Tx) ) {
   SampledPt t_info;
   for ( int i = 0; i < MARCH_REPS; ++ i ) {
     t_info = Map(avoid, ray.origin + ray.dir*distance, si, Tx);
[…]

### 2017-10-22 raytracekernel.c

@@ -164,11 +164,11 @@ SampledPt March ( int avoid, Ray ray, SCENE_T(si, Tx) ) {
 
 float Visibility_Ray(float3 orig, float3 other, float sphere_radius,
                      SCENE_T(si, Tx)) {
-  float theoretical = Distance(orig, other) - sphere_radius;
+  float theoretical = Distance(orig, other) - sphere_radius*2.0f;
   float3 dir = normalize(other - orig);
-  orig += dir*(MARCH_ACC*2.0f + 0.2f);
+  orig += dir*(MARCH_ACC*2.0f); // get out of our own intersection
   SampledPt ptinfo = March(-1, (Ray){orig, dir}, si, Tx);
-  float actual = ptinfo.dist + MARCH_ACC + 0.4f;
+  float actual = ptinfo.dist + MARCH_ACC*1.2f; // distance + leeway
   return 1.0f*(actual >= theoretical);
 }
 

### 2017-11-21 dtoadqkernel.c
added dtq brdf

@@ -29,6 +29,7 @@
 #define Spectrum float3
 
 __constant float MARCH_ACC = //%MARCH_ACC.0f/1000.0f;
+__constant int   DO_NAVIGATION = 1;
 __constant int   M_diffuse      = 1, M_glossy       = 2,
                  M_specular     = 3, M_transmittive = 4;
 // -----------------------------------------------------------------------------
@@ -51,14 +52,11 @@ typedef struct T_Camera {
 typedef struct T_Material {
   // colour [set to (-1.0, -1.0, -1.0) to have map override it]
   float3 albedo;
-  // pdf
+  // sampling strategy
   float diffuse, specular, glossy, glossy_lobe;
-  float transmittive, ior;
+  float transmittive;
   // PBR material
-  float pbr_roughness, pbr_metallic, pbr_fresnel;
-  // disney material
-  float d_subsurface, d_specular_tint, d_anisotropic, d_sheen, d_sheen_tint,
-        d_clearcoat, d_clearcoat_gloss, d_specular;
+  float roughness, metallic, fresnel, subsurface, anisotropic;
 } Material;
 
 typedef struct T_Emitter {
@@ -97,10 +95,15 @@ SampledPt SampledPt_From_Origin(float3 origin) {
 }
 
 typedef struct T_Vertex {
-  float3 origin, albedo, normal, pdf_forward, pdf_backward;
+  float3 origin, albedo, normal;
+  float pdf_fwd, pdf_bwd;
   Material* material;
   int M_ID;
 } Vertex;
+
+float Calc_Prob(Vertex* t) {
+  return t->pdf_bwd/t->pdf_fwd;
+}
 
 typedef struct T_Subpath {
   Vertex vertices[MAX_DEPTH];
@@ -259,8 +262,15 @@ float3 refract(float3 V, float3 N, float refraction) {
   return (refraction*V) + (refraction*cosI - sqrt(cosT))*N;
 }
 
-float3 To_Cartesian ( float sin_theta, float cos_theta, float phi ) {
+float3 To_Cartesian ( float cos_theta, float phi ) {
+  float sin_theta = sqrt(fmax(0.0f, 1.0f - cos_theta));
   return (float3)(cos(phi)*sin_theta, sin(phi)*sin_theta, cos_theta);
+}
+
[…]

### 2017-11-21 raycastkernel.c

@@ -105,7 +105,7 @@ SampledPt Map ( int a, float3 origin, SCENE_T(si, Tx) ) {
   for ( int i = 0; i != EMITTER_AMT; ++ i ) {
     Emitter e = REmission(i, si->debug_values, si->time);
     float dist = sdSphere(origin - e.origin, e.radius);
-    MapUnionG(a, &res, dist, -100 - i, (float3)(1.0f, 0.9f, 0.8f)*e.emission);
+    MapUnionG(a, &res, dist, -100 - i, e.emission);
   }
 
   return res;

### 2017-11-24 documentation/_region_.tex
almost done with bpt just need mis
added better subsurface to dtq, fixed BDPT bugs, added S=0 strategy

@@
   \section{SDFs}
+  In order to compute a normal of an SDF, you must perform a gradient approximation: the central distance onf the SDF at
+  P. For example, the difference quotient:
+  \[\frac{(f(x+h) - f(x))}{h}\]
   \begin{align*}
@@
   \section{Bidirectional Path Tracing}
-    All the materials in here, about path tracing, have come from Veach's PhD
-    thesis. So, in a way, DTOADQ, at this stage, is just an implementation of
-    BDPT as described in veach's thesis to the GPU, along with pulling from
-    other resources such as Physically Based Render [PBR] 3rd ed. In order
-    to understand the mathematics for BDPT, it must be worked up starting from
-    light transport. So this section is about transforming the algorithms and
-    mathematical models into something easily and efficiently implementable for
-    the GPU, even if some of it is a rewrite of the mathematical model so that a
-    programmer could understand it. In order to compute a normal of an SDF, you
-    must perform a gradient approximation: the central distance onf the SDF at
-    P. For example, the difference quotient:
-      \[\frac{(f(x+h) - f(x))}{h}\]
+    The main foundation of the mathematics and theory have come from many sources, from Veach's PhD thesis to PBR 3rd ed.
+and various open source path tracing kernels such as NanoRT, Embree, etc.
+However, the implementation differs from these in that the algorithm is optimized for the GPU.
+This section is about transforming the algorithms and mathematical models from Veach's thesis into something easily and efficiently implementable the GPU.
 
@@
   \begin{align}  
- \includegraphics{bdpt_unweighted}
   \end{align}
@@
 
-\message{ !name(paper.tex) !offset(-461) }
+\message{ !name(paper.tex) !offset(-459) }

### 2017-11-24 documentation/paper.tex
added better subsurface to dtq, fixed BDPT bugs, added S=0 strategy

@@
 
-\title{DTOADQ}
+\title{DTOADRe}
 

### 2017-11-24 dtoadqkernel.c

@@ -95,7 +95,7 @@ SampledPt SampledPt_From_Origin(float3 origin) {
 }
 
 typedef struct T_Vertex {
-  float3 origin, albedo, normal;
+  float3 origin, radiance, normal, mat_col;
   float pdf_fwd, pdf_bwd;
   Material* material;
   int M_ID;
@@ -176,6 +176,13 @@ float Power_Heuristic ( float fn, float fpdf, float gn, float gpdf ) {
         g = sqr(gn*gpdf);
   return f/(f + g);
 }
+
+float3 Float3_Max(float3 t, float g) {
+  return (float3)(fmax(t.x, g), fmax(t.y, g), fmax(t.z, g));
+}
+float3 Float3_Min(float3 t, float g) {
+return (float3)(fmin(t.x, g), fmin(t.y, g), fmin(t.z, g));
+}
 // -----------------------------------------------------------------------------
 // --------------- MAP GEOMETRY FUNCTIONS --------------------------------------
 
@@ -249,7 +256,7 @@ float3 Normal ( float3 p, SCENE_T(s, t)) {
 }
 
 float3 RColour ( float3 pt_colour, Material* m ) {
-  return (m->albedo.x < 0.0f) ? pt_colour : m->albedo;
+  return (pt_colour.x >= 0.0f) ? pt_colour : m->albedo;
 }
 
 float3 reflect ( float3 V, float3 N ) {
@@ -273,34 +280,15 @@ float3 Binormal ( float3 N ) {
   return normalize(cross(N, axis));
 }
 
-// from PBRT 3rd ed
-float2 Concentric_Sample_Disk(SceneInfo* si) {
-  float2 u = Sample_Uniform2(si);
-  // maps to [-1, 1]^2
-  float2 offset = 2.0f * u - (float2)(1.0f, 1.0f);
-  if ( offset.x == 0.0f && offset.y == 0.0f )
-    return (float2)(0.0f);
-
-  float theta, r;
-  if ( fabs(offset.x) > fabs(offset.y) ) {
-    r = offset.x;
-    theta = (PI/4.0f) * (offset.y/offset.x);
-  } else {
-    r = offset.y;
-    theta = (PI/2.0f) * (offset.x/offset.y);
-  }
-  return r * (float2)(cos(theta), sin(theta));
-}
-
-// from PBRT 3rd ed
+// Most sample functions from embree
[…]

### 2017-11-24 raycastkernel.c

@@ -32,7 +32,13 @@ typedef struct T_Camera {
 } Camera;
 
 typedef struct T_Material {
-  float diffuse, specular, glossy, retroreflective, transmittive;
+  // colour [set to (-1.0, -1.0, -1.0) to have map override it]
+  float3 albedo;
+  // sampling strategy
+  float diffuse, specular, glossy, glossy_lobe;
+  float transmittive;
+  // PBR material
+  float roughness, metallic, fresnel, subsurface, anisotropic;
 } Material;
 
 typedef struct T_Emitter {
@@ -134,14 +140,8 @@ float Distance(float3 u, float3 v) {
   return sqrt(x*x + y*y + z*z);
 }
 
-float Visibility(float3 orig, float3 other, float sphrad, SceneInfo* si,
-                 TEXTURE_T Tx) {
-  float theoretical = Distance(orig, other);
-  float3 dir = normalize(other - orig);
-  orig += dir*(0.01f);
-  SampledPt ptinfo = March(-1, (Ray){orig, dir}, si, Tx);
-  float actual = ptinfo.dist + MARCH_ACC + 0.02f;
-  return 1.0f*(actual >= theoretical);
+float3 RColour ( float3 pt_colour, Material* m ) {
+  return (pt_colour.x >= 0.0f) ? pt_colour : m->albedo;
 }
 
 //----POSTPROCESS---
@@ -201,12 +201,18 @@ __kernel void DTOADQ_Kernel (
 
   Ray ray = Camera_Ray(&camera);
   SampledPt pt = March(-1, ray, &scene_info, textures);
-  pt.colour += pt.dist/MARCH_DIST;
-  float3 O = ray.origin + pt.dist*ray.dir;
-  for ( int i = 0; i != EMITTER_AMT; ++ i ) {
-    Emitter emit = REmission(i, scene_info.debug_values, scene_info.time);
-    float res = Visibility(O, emit.origin, emit.radius, &scene_info, textures);
-    pt.colour *= (res > 0.0f ? 1.0f : 0.2f);
+  if ( pt.dist >= 0 ) {
+    float3 O = ray.origin + pt.dist*ray.dir;
+    for ( int i = 0; i != EMITTER_AMT; ++ i ) {
+      Emitter emit = REmission(i, scene_info.debug_values, scene_info.time);
+      Material m = materials[pt.material_index];
+      float3 col = RColour(pt.colour, &m);
[…]

### 2017-11-24 raytracekernel.c

@@ -37,14 +37,11 @@ typedef struct T_Camera {
 typedef struct T_Material {
   // colour [set to (-1.0, -1.0, -1.0) to have map override it]
   float3 albedo;
-  // pdf
+  // sampling strategy
   float diffuse, specular, glossy, glossy_lobe;
-  float transmittive, ior;
+  float transmittive;
   // PBR material
-  float pbr_roughness, pbr_metallic, pbr_fresnel;
-  // disney material
-  float d_subsurface, d_specular_tint, d_anisotropic, d_sheen, d_sheen_tint,
-    d_clearcoat, d_clearcoat_gloss, d_specular;
+  float roughness, metallic, fresnel, subsurface, anisotropic;
 } Material;
 
 typedef struct T_Emitter {
@@ -67,8 +64,11 @@ SceneInfo New_SceneInfo(float time, Material* materials,
   return si;
 }
 
-__constant float PI  = 3.141592654f;
-__constant float TAU = 6.283185307f;
+__constant float PI   = 3.141592653589793f;
+__constant float IPI  = 0.318309886183791f;
+__constant float IPI2 = 0.159154943091895f;
+__constant float TAU  = 6.283185307179586f;
+__constant float ITAU = 0.159154943091895f;
 // -----------------------------------------------------------------------------
 // --------------- GENERAL STRUCTS ---------------------------------------------
 typedef struct T_Ray {
@@ -186,103 +186,78 @@ float3 Binormal ( float3 N ) {
                                     (float3)(0.0f, 1.0f, 0.0f));
   return normalize(cross(N, axis));
 }
+
 float Schlick_Fresnel ( float u ) {
-  float m = clamp(1.0f - u, 0.0f, 1.0f);
-  float m2 = m*m;
-  return m2*m2*m;
+  float f = clamp(1.0f - u, 0.0f, 1.0f);
+  float f2 = f*f;
+  return f2*f2*f; // f^5
 }
-float GTR1 ( float cosN_H, float a ) {
-  if ( a >= 1.0f ) return 1.0f/PI;
-  float a2 = a*a;
-  float t = 1.0f + (a2-1.0f) * sqr(cosN_H);
-  return (a2 - 1.0f)/(PI*log(a2)*t);
+
+float Smith_G_GGX_Correlated ( float L, float R, float a ) {
+  return L * sqrt(R - a*sqr(R) + a);
 }
[…]

### 2017-11-27 dtoadqkernel.c
adding transmittive

@@ -24,8 +24,6 @@
 
 // math macros
 #define SQR(T) ((T)*(T))
-
-//
 #define Spectrum float3
 
 __constant float MARCH_ACC = //%MARCH_ACC.0f/1000.0f;
@@ -54,7 +52,7 @@ typedef struct T_Material {
   float3 albedo;
   // sampling strategy
   float diffuse, specular, glossy, glossy_lobe;
-  float transmittive;
+  float transmittive, ior;
   // PBR material
   float roughness, metallic, fresnel, subsurface, anisotropic;
 } Material;
@@ -228,7 +226,7 @@ SampledPt Map ( int a, float3 origin, SCENE_T(si, Tx))  {
 // -----------------------------------------------------------------------------
 // --------------- RAYMARCHING      --------------------------------------------
 SampledPt March ( int avoid, Ray ray, SCENE_T(si, Tx)) {
-  float distance = 0.001f;
+  float distance = 0.01f;
   SampledPt t_info;
   for ( int i = 0; i < MARCH_REPS; ++ i ) {
     t_info = Map(avoid, ray.origin + ray.dir*distance, si, Tx);
@@ -263,10 +261,10 @@ float3 reflect ( float3 V, float3 N ) {
   return V - 2.0f*dot(V, N)*N;
 }
 
-float3 refract(float3 V, float3 N, float refraction) {
-  float cosI = -dot(N, V);
-  float cosT = 1.0f - refraction*refraction*(1.0f - cosI*cosI);
-  return (refraction*V) + (refraction*cosI - sqrt(cosT))*N;
+float3 refract(float3 I, float3 N, float ior) {
+  float cos_NI = dot(N, I);
+  float k = 1.0f - SQR(ior)*(1.0f - sqr(cos_NI));
+  return k < 0.0f ? (float3)(0.0f) : ior*I - (ior*cos_NI + sqrt(k))*N;
 }
 
 float3 To_Cartesian ( float cos_theta, float phi ) {
@@ -361,19 +359,21 @@ float3 BRDF_Specular_Sample ( float3 wi, float3 N, float* pdf ) {
   *pdf = 1.0f;
   return reflect(wi, N);
 }
+// Misnomer, it should be either BSDF or BTDF ; TODO make all pdf/sample/etc bsdf
+float3 BRDF_Transmittive_Sample ( float3 wi, float3 N, float* pdf, float ior ) {
+  *pdf = 1.0f;
[…]

### 2017-11-27 raycastkernel.c

@@ -36,7 +36,7 @@ typedef struct T_Material {
   float3 albedo;
   // sampling strategy
   float diffuse, specular, glossy, glossy_lobe;
-  float transmittive;
+  float transmittive, ior;
   // PBR material
   float roughness, metallic, fresnel, subsurface, anisotropic;
 } Material;

### 2017-11-27 raytracekernel.c

@@ -17,6 +17,10 @@
 #define writefloat3(X) if (Is_Debug()){printf(#X " %f, %f, %f\n",    DFLT3(X));}
 #define writefloat4(X) if (Is_Debug()){printf(#X " %f, %f, %f, %f\n",DFLT4(X));}
 
+// math macros
+#define SQR(T) ((T)*(T))
+#define Spectrum float3
+
 __constant float MARCH_ACC = //%MARCH_ACC.0f/1000.0f;
 // -----------------------------------------------------------------------------
 // --------------- DEBUG -------------------------------------------------------
@@ -39,7 +43,7 @@ typedef struct T_Material {
   float3 albedo;
   // sampling strategy
   float diffuse, specular, glossy, glossy_lobe;
-  float transmittive;
+  float transmittive, ior;
   // PBR material
   float roughness, metallic, fresnel, subsurface, anisotropic;
 } Material;
@@ -82,6 +86,9 @@ typedef struct T_SampledPt {
 } SampledPt;
 // -----------------------------------------------------------------------------
 // --------------- GENERAL UTILITIES      --------------------------------------
+float To_IOR ( float ior ) {
+  return 0.0f + 2.5f*ior;
+}
 float Distance(float3 u, float3 v) {
   float x = u.x-v.x, y = u.y-v.y, z = u.z-v.z;
   return sqrt(x*x + y*y + z*z);
@@ -143,10 +150,10 @@ float3 reflect ( float3 V, float3 N ) {
   return V - 2.0f*dot(V, N)*N;
 }
 
-float3 refract(float3 V, float3 N, float refraction) {
-  float cosI = -dot(N, V);
-  float cosT = 1.0f - refraction*refraction*(1.0f - cosI*cosI);
-  return (refraction*V) + (refraction*cosI - sqrt(cosT))*N;
+float3 refract(float3 I, float3 N, float ior) {
+  float cos_NI = dot(N, I);
+  float k = 1.0f - SQR(ior)*(1.0f - sqr(cos_NI));
+  return k < 0.0f ? (float3)(0.0f) : ior*I - (ior*cos_NI + sqrt(k))*N;
 }
 
 // -----------------------------------------------------------------------------
@@ -197,7 +204,7 @@ float Smith_G_GGX_Correlated ( float L, float R, float a ) {
   return L * sqrt(R - a*sqr(R) + a);
 }
 
[…]

### 2017-11-27 testdocs/test.tex
doicumentation

--- 2017-11-24 documentation/paper.tex
@@
 \usepackage{graphicx}
-\graphicspath{ {../screenshots/} }
+\graphicspath{ {images/} }
 \usepackage[margin=1.15in]{geometry}
@@
 
-\title{DTOADRe}
+\title{DTOADQ}
 
@@
 
-  \section{SDFs}
+  \section{Signed Distance Fields}
+  Signed Distance Fields (SDFs) are a mathematical tool to represent a 3D primitive by describing its surface via a distance field.
+Given an arbitrary point in space, the SDF of a primitive is the distance to its nearest point/surface.
+For example, the SDF of a sphere is $\Vert\:O\:\Vert - R$.
+With SDFs, it's possible to march a ray along the scene until it hits the surface of an SDF within a given bound, using the returned SDF distance as the distance to march.
+Some example SDFs are shown below:    \begin{align*}
+  S(P, R) &= \Vert\:P\:\Vert - R & \text{Sphere with Radius}\\
+  S(P, N) &= P \cdot N  & \text{Plane with Normal}\\
+  S(P, H, R) &= <\Vert\:P_{x, z}\:\Vert - R, \: P_{y}> - H & \text{Torus with Radius and Hole}\\
+  \end{align*}
+
+  The biggest issue with SDFs is Lipschitz continuity.
+The SDFs themselves have a distance gradient, that is, their bound.
+The bounds of a SDF is expected to be at most 1 in order to remain Lipschitz continuous, as otherwise the SDF will appear deformed as shown below.
+
+  \includegraphics[]{lipschittzbound}
+
+  This is a larger problem in Light Transport than it is in raytracing because small discontinuities can impact the lighting of an entire scene.
+And to add to that, most interesting operations (such as the meta balls shown in the image above) will violate Lipschitz.
+The easiest way to fix this is to multiply the distance field by a constant to stay below 1, but this is less of a fix and more of a hack.
+
   In order to compute a normal of an SDF, you must perform a gradient approximation: the central distance onf the SDF at
@@
   \[\frac{(f(x+h) - f(x))}{h}\]
-  \begin{align*}
-    S(P, R) &= P - R^2   & \text{SDF of a sphere Radius  }\\
-    S(P, N) &= P \cdot N  & \text{SDF of a plane at Normal}\\
-    S(P, H, R) &= <|P_{x, z}| - R, \: P_{y}> - H & \text{SDF of a torus with
-                                                  Radius and Hole}\\
-  \end{align*}
+
   \section{OpenCL}
+	OpenCL, and GPU programming in general, is where I am least knowledgeable or interested.
+Well, OpenCL is easy to work with, and I wrote my own OpenCL module to handle OCL in a rather convenient method, for example, the DTOADQ kernel call looks as such:
+	\begin{lstlisting}
+	shared_info.image.Lock(); // Acquire OCL Mem from OGL image object
+	ocl.Run(
+	  ocl.CLStoreMem(shared_info.rw_image),
+	  shared_info.image.RWrite(),
+	  ocl.CLStoreMem(shared_info.image_metadata),
+	  shared_info.camera,
+	  shared_info.timer,
+	  ocl.CLPredefinedMem(imgs),
+	  ocl.CLStoreMem(shared_info.ocl_material),
+	  debug_values,
+	  ocl.CLStoreMem(shared_info.rng_states),
+	  // kernel size
+	  shared_info.image.x, shared_info.image.y
+	);
+	shared_info.image.Unlock();
+
+	\end{lstlisting}
+
+	however the problem comes with dealing with the intrinsics of GPU; optimizing code for the GPU and knowing how memory and threads/workers are handled.
+
+  	One benefit of OpenCL is its ease to debug in comparison to OpenGL.
+One immediate benefit is printf, which surprisingly even works on the GPU.
+To make printf easily readable, I only use print on the center pixel of the image.
+Even so, due to the complexity of light transport, the vast majority of time spent on this project went directly towards debugging the DTOADQ kernel.
+
+  	OpenCL also makes it relatively easy to send and receive data between the host and GPU.
+In comparison to OpenGL, this is a big win, however, with OpenGL4.3's Shader Storage Buffer Object, this is now no longer a selling point.
+
+  	OpenCL also has huge drawbacks.
+The biggest problem is getting the results to be rendered in OpenGL.
+TODO.
+Also, there are minor annoyances, like the lack of a matrix data type, no function overloads, long compile times, or variadic functions.
+There are also annoyances with GPU programming in general.
+No dedicated random function.
+Though, possibly, long compile times is my fault for using a mega-kernel instead of splitting it into many small kernels.
+
+  	It seems that the GPU is not for me or DTOADQ and something I do not enjoy to use nor am good at.
+However, for this project, I had to deal with it.
+To overcome the lack of a random function, of which high quality uniform random generation is necessary for Monte Carlo simulations, I used the MWC64X algorithm.
+While there are better algorithms, they fail on branches which makes it useless for DTOADQ.
+It seemed to work well enough.
+The downside is that each worker/pixel needs its own random state, preserved across multiple kernel calls.
+There are indefinitely better solutions.
   \section{Light Transport Introduction}
@@
     biggest conceptual barrier is that, for example, given a solid angle in a
-    hemisphere $\Omega$, the probability for a random variable X to equal
-    $\omega$ is 0. But this isn't what's happening here, this is all about the
-    probability that X could potentially equal $\omega$ when compared to other values
-    in $\Omega$. It's all about comparing the probabilities of two potential
-    $\omega$s with each other.
+    hemisphere $\Omega$, the probability for a random uniform $X$ to equal
+    $\omega$ is 0. That's not what the PDF calculates, instead, it's the
+    probability that a random uniform $X$ could potentially equal $\omega$ when compared to other $\omega\:$s.
+In other words, it's about comparing the probability of $X$ to equal $\omega$ when compared to another theoretical z$\omega$.
 
-    I still have a hard time conceptualizing these things in my head, but the
-    good thing is, is that calculating these are relatively easy. There's a lot
-    that could be talked about, but the important thing is that I've found a
+    The good thing is that calculating these are relatively easy. There's a
     reliable three-step process. The first is to
@@
     set of all possible values is correlated to surface area:
-    $\Omega = [0,\:2.0f*\tau*R^2]$.
-    Thus for a given constant $x$, and a uniformly random $X$,
+    $\Omega = [0,\:2.0f*\tau*R^2]$. Thus,
     \begin{align}
-      CDF(x) = Pr(X \leq x) = \frac{x}{2.0f*\tau*R^2}\\
-      p(x) = \frac{dCDF(x)}{dx} = \frac{1.0f}{2.0f*\tau*R^2}\\
-      \int_0^{\Omega}p(x)dx = 1.0f
+      CDF(x) = Pr(X \leq x) = \frac{x}{2.0*\tau*R^2}\\
+      p(x) = \frac{dCDF(x)}{dx} = \frac{1.0}{2.0*\tau*R^2}\\
+      \int_0^{\Omega}p(x)dx = 1.0
     \end{align}
@@
     \begin{lstlisting}
-      float Light_PDF ( float3 P, Emitter* m ) {
-        float rad_sqr = m->radius*m->radius;
-        return 1.0f/(2.0f*TAU*rad_sqr);
+      float Light_PDF ( Emitter* m ) {
+        return 1.0f/(2.0f*TAU*SQR(m->radius));
       }
@@
     \begin{cases}
-      L_e(z_{-2} \rightarrow z_{-1}), &s = 0, t > 0,\\
+      L_e(z_{-2} \rightarrow z_{-1}), &s = 0, t > 1,\\
       f_s(y_{-2} \rightarrow y_{-1} \rightarrow z_{-1}) G(y_{-1} \leftrightarrow
-      z_{-1}) f_s(y_{-1} \rightarrow z_{-1} \rightarrow z_{-2}), &s > 0, t > 0
+      z_{-1}) f_s(y_{-1} \rightarrow z_{-1} \rightarrow z_{-2}), &s > 0, t > 1
     \end{cases}
   \end{align}
-    One question to be made is if the strategy $s = 0, t > 0$ is worthwhile to
-    compute; the most noticeable effect are for caustics, light sources in
-    direct view of an emitter should be emitted from animations in this current
-    state of DTOADQ. Another idea to consider is that under real physics
-    circumstances, the photons are still allowed to bounce after hitting a
-    source of light, but the contribution from these cases are negligible, and
-    so most likely not worthwhile to implement. Anyways, the $s = 0$
-    strategy is a worthwhile  calculation under these circumstances
+    Is the $s = 0, t > 1$
+    strategy worthwhile to compute? Those familiar with next event
+    estimation would know that there are situations in which this is beneficial.
+    Take, for example, a glossy surface with a small solid angle subtending to
+    the surface of a large light source. The $s=1$ sample strategy in this case
+    would only contribute when the $L0$ vertex is within that solid angle.
+    The effective strategy to sample in this case is $s = 0$.
+    In code, $\alpha_i^E\:c_{s,t}\:\alpha_0^L$ looks like (MIS not shown):
+    \begin{lstlisting}
+     float3 wi = normalize(E1.origin - E0.origin),
+            wo = normalize(L1.origin - E1.origin);
+     float3 albedo = _BRDF_F(wi, E1.normal, wo, E1.material, E1.mat_col);
+     albedo /= BSDF_PDF(wo, wi, &E1);
+     Spectrum contribution = E1.radiance * albedo;
+     contribution *= Geometric_Term(&E1, &L1, si, Tx);
+     contribution *= tlight.emission/Light_PDF(&tlight);
+    \end{lstlisting}
+
+    Another question to consider is if the path should be able to reflect on the
+    light source. While under real circumstances, this is allowed, the
+    contribution is negligible and dismissed in DTOADQ. Also, there are no cases for $s = s, t = 0|1$ as
+    the lens does not have a physical existence in the SDF map.
+The $t = 0$ case, where a photon from the light source hits the lens directly, is rather pointless anyways, as a path generated by light-tracing would have to be able to contribute to any pixel, which is not feasible on a GPU.
+The case of $t = 1$, where the light path is connected to a position on the lens, is important for effects such as depth of field and aperture.
+The only problem is due to time constraints, this has not been implemented yet.
 
@@
     cancelled out for the entire equation, except for the connection term $c$.
-    The expansion is not shown as it is very lengthy, and unnecessary. Paths
-    $\bar{y}_{0} \ldots \bar{y}_s$ and $\bar{z}_0 \ldots \bar{z}_t$ being
-    visible is intuitively obvious as the paths have been generated using the
-    same technique used to check for visibility; raymarching. Specifically, the
-    definition of $V$ is
+    This is intuitive as the paths
+    $\bar{y}_{0} \ldots \bar{y}_s$ and $\bar{z}_0 \ldots \bar{z}_t$ had just been generated via raymarching.
+    Specifically, the definition of $V$ is
   \begin{align}
@@
     the problem is that the $\bar{y}$ path is generated in perspective of the
-    light source, so we have to call the BSDF in perspective of the camera; the
-    path is flipped. Handling $\bar{y}$ is significantly easier than $\bar{z}$,
-    as $\bar{y}$ is precomputed. For $\bar{z}$, the next vertex $z_{+1}$
-    ($\omega_o$ in the perspective of SA), is unknown. For example $\alpha^L_1$
-    in OpenCL looks like:
+    light source, so there has to be a bit of caution in terms of perspective in when generating the PDFs.
 
-    \begin{lstlisting}
-      Vertex* V0 = light_path.vertices - 1, * V1 = light_path.vertices;
-      float sigma_pdf = Light_PDF(V1.origin, &light);
-      float g = Geometric_Term(V0.origin, V1.origin, V0.normal, V1.normal);
-      float light_pdf = sigma_pdf * g;
-      float Le = light.colour;
-      light_contrib[1] = Le / light_pdf;
-    \end{lstlisting}
-
-    For $c_{s, t}$, there is no cases for $s = s, t = 0$ as
-    the lens does not have a physical existence in the SDF map. $p_A$ is
+     $p_A$ is
     the PDF w.r.t. area, for the light path; we only need to take the PDF of the
@@
   worth the additional computation time that BDPT requires, as shown below.
-  \begin{align}  
-  \end{align}
+  	 To calculate the PDF of 
   In order to get good
@@
 
+  \section{DTOADQ BRDF}
+  DTOADQ supports a "universal" physically-based sampling and spectrum BRDF.
+The spectrum BRDF is a combination of many different modified BRDF functions into one.
+There is a microfacet, subsurface and diffusive component in the DTOADQ spectrum BRDF, which are all combined together, with many variables to tune these components.
+There is diffusive, glossy, specular and transmittive sampling in the BRDF importance sample, with variables to allow specific control over which strategies to choose and how they are combined together.
+Regardless of how the BRDF is sampled, they use the same spectrum BRDF.
+
+  In order to develop these BRDFs, I wrote an SDF model based off the mitsuba model and wrote an IBL path-traced shader for quick prototyping; shadertoy.com/view/XtSyDm .
+The main reason to not use DTOADQ was the lack of cubemap support, which would have taken too long to implement.
+
+  The Microfacet model consists of a fresnel, distribution and geometric term.
+It's important to note that these have all been modified to use the half-vector $H = \Vert\:L+V\:\Vert$.
+This gives better visuals and allows for a specific energy conservation model.
+Anyways, the fresnel term approximates the fresnel factor, it's based off both Schlick's approximation and Disney's diffusive schlick fresnel, with modifications to use a $metallic$ term as well as an albedo;
+  \begin{lstlisting}
+const float3 Fresnel_L = Schlick_Fresnel(cos_NL),
+             Fresnel_V = Schlick_Fresnel(cos_NV);
+const float F0 = m->fresnel * m->metallic,
+Fresnel_diffuse_90 = F0 * SQR(cos_HL);
+microfacet *= (1.0f - F0) * diffusive_albedo +
+    mix(1.0f, Fresnel_diffuse_90, Fresnel_L) *
+    mix(1.0f, Fresnel_diffuse_90, Fresnel_V);
+  \end{lstlisting}
+  \includegraphics[scale=0.20]{fresnel}
+
+  The Geometric term is used to control how microfacets 'shadow' each other.
+The DTQ BRDF uses Heits' [2014] SmithGGXCorrelated modified to use a half vector, combined with an anisotropic term based off the GGX Anisotropic model:
+  \begin{lstlisting}
+  const float Param  = sqr(0.5f + sqr(m->roughness)),
+  Aspect = sqrt(1.0f - m->anisotropic*0.9f),
+  Ax     = Param/Aspect, Ay = Param*Aspect,
+  GGX_NV = Smith_G_GGX_Correlated(cos_HL, cos_NV, Ax),
+  GGX_HL = Smith_G_GGX_Correlated(cos_NV, cos_HL, Ay);
+  microfacet *= 0.5f / (GGX_NV*Ax + GGX_HL*Ay);
+  \end{lstlisting}
+  \includegraphics[scale=0.20]{geometric}
+
+  And finally, the Distribution term controls how microfacets are spread around the normal.
+It's based off the Hyper Cauchy with a roughness and metallic (shape) parameters [9].
+The latter of which should most likely be its own parameter in the future.
+The more readable mathematical model is shown:
+
+  \begin{align}
+  \frac{(M - 1)\:(\sqrt{2})^{2\:S\:-\:2}}
+  	    {\pi\:R^{2}\:cos^{4}\theta\:(2 +
+  	      \frac{tan^{2}\theta}{M^2})^M}
+  \end{align}
+  \includegraphics[scale=0.20]{distribution}
+
+  These combine to form the microfacet under the half vector microfacet energy conservation model [Edwards et al. 2006].
+  \[
+  	\frac{F*G*D}{4 \: (H \cdot V) \: ( (N \cdot L) \vee (N \cdot V) )}
+  \]
+
+  The diffusive component is a combination of both the subsurface and lambertian component.
+The lambertian is simply $\alpha^{0.5} * \frac{1}{\pi}$, and the subsurface component is based off the Henyey-Greenstein formula [Henyey], modified for half-vec:
+  \begin{align}
+  	\frac{m(1 - r^2)}{4\pi}\:\frac{1}{(1 + r^2 - 2\:r\:
+  		(H \cdot V))^{3/2}}
+  \end{align}
+  where m is metallic and r is roughness. This is then combined to the diffusive component of the BRDF using Fresnel factor like so:
+  \begin{lstlisting}
+const float R = (-0.3f + 1.3f*m->roughness)*cos_HV,
+            M = (1.0f - m->metallic)*5.0f;
+const float Rr_term = M * (1.0f - sqr(R))*IPI *
+                      (1.0f/(pow(1.0f + sqr(R) - 2.0*R*cos_HL, 3.0f/2.0f)));
+const float3 Retro_reflection = diffusive_albedo * Rr_term *
+                  (Fresnel_L + Fresnel_V + (Fresnel_L*Fresnel_V*(Rr_term)));
+diffusive_albedo = mix(diffusive_albedo, Retro_reflection, m->subsurface);
+
+  \end{lstlisting}
+  TODO: With 0.95 to exaggerate
+
+  The microfacet and diffusive components are then added together.
+
   \section{Video/Image Emitter}
@@
   \end{align*}
+  [9]
+  Samuel D. Butler, Experimental and Theoretical Basis for a closed-Form Spectral BRDF model
+
+  [Henyey]
+   Diffuse Radation in the Galaxy, 1942. L, G. Henyey and J. L. Greenstein. 
 \end{document}

### 2017-11-27 testdocs/test.tex
doicumentation

@@
   \[\frac{(f(x+h) - f(x))}{h}\]
-
+  The most obvious way to calculate this is the six-point gradient; where for
+each axis you compute $f(x+h) - f(x-h)$, and normalize the result. This
+returns a vector pointing in the direction where the map's SDF changes the
+most: the normal. A four point gradient is possible; as long as each each
+point of the gradient is multiplied by $\epsilon$ before normalizing. The
+speedup from six to four is worthwhile, as computing each gradient is
+expensive as it involves a sdf map call. There are more utilities with
+gradients not yet currently provided in DTOADQ but are worth mentioning, a
+few are: an additional normal call using an alternative (higher-quality) SDF
+map can be used to compute a 'normal' bumpmap, and analyzing the gradient by
+hand allows a cheap normal computation of primitives such as a spheres.
   \section{OpenCL}
@@
     Monte Carlo Sampling. The PDF states the probability of a random variable to
-    take the form of a given input. Its usefulness comes when contributions from
-    many Monte Carlo samples have a different weight. That is, one sample may
+    take the form of a given input. Its usefulness comes when a Monte Carlo sample has a smaller or higher total contribution compared to other samples. That is, one sample may
     have a higher probability distribution, and thus a higher contribution to
@@
     The PDF applies to everything from path contributions, solid angles of a
-    BRDF, uniformly sampling a sphere, etc. I had a difficult time understanding
-    these, and more importantly, how to calculate them for a given formula. The
-    biggest conceptual barrier is that, for example, given a solid angle in a
-    hemisphere $\Omega$, the probability for a random uniform $X$ to equal
-    $\omega$ is 0. That's not what the PDF calculates, instead, it's the
+    BRDF, uniformly sampling a sphere, etc.
+
+    The
+    biggest conceptual barrier is that, for example, given a solid angle in a hemisphere $\Omega$, the probability for a random uniform $X$ to equal $\omega$ is 0.
+That's not what the PDF calculates, instead, it's the
     probability that a random uniform $X$ could potentially equal $\omega$ when compared to other $\omega\:$s.
-In other words, it's about comparing the probability of $X$ to equal $\omega$ when compared to another theoretical z$\omega$.
 
-    The good thing is that calculating these are relatively easy. There's a
-    reliable three-step process. The first is to
+    The good thing is that calculating these are relatively easy. The first step is to
     find the cumulative distribution function $CDF = Pr(X \leq x)$. Where $Pr$
     is the probability, $X$ is a uniformly random variable, and $x$ is a known
-    constant. In other words, this gives the probability that, for a formula
+    constant. This gives the probability that, for a formula
     $P$, where $X \in \Omega$, and $\Omega$ is the set of all possible values
     for $CDF$,
-    that $X \leq x$. The second step is to take the derivative of this, so for
-    PDF $p$
-    \begin{align}
-      p(x) = \frac{dCDF(x)}{dx}
-    \end{align}
-    And the final step is to verify that the PDF is correct by integrating it
+    that $X \leq x$. Then the PDF $p$ is defined by the derivative of CDF. To verify that the PDF is correct, integrate it
     over $\Omega$, this final value should equal 1. This makes sense intuitively
@@
     For an example, how I calculated the PDF of an area light source
-      $L^0_e(v) = L_o(P, \omega_o)$ 
+      $L^0_e(v) = L_o(P,\:\omega_o)$ 
     is quite simple. Firstly,the area light is defined as a sphere, and thus our
@@
     \end{lstlisting}
-    The more interesting aspect, in this example, is actually generating the
-    uniformly distributed samples of a sphere. The naive method gives incorrect
-    distributions.TODO {page 774}
 
-    The most obvious way to calculate this is the six-point gradient; where for
-    each axis you compute $f(x+h) - f(x-h)$, and normalize the result. This
-    returns a vector pointing in the direction where the map's SDF changes the
-    most: the normal. A four point gradient is possible; as long as each each
-    point of the gradient is multiplied by $\epsilon$ before normalizing. The
-    speedup from six to four is worthwhile, as computing each gradient is
-    expensive as it involves a sdf map call. There are more utilities with
-    gradients not yet currently provided in DTOADQ but are worth mentioning, a
-    few are: an additional normal call using an alternative (higher-quality) SDF
-    map can be used to compute a 'normal' bumpmap, and analyzing the gradient by
-    hand allows a cheap normal computation of primitives such as a spheres.
+  The rendering equation is:
+  \begin{align}
+    L_o(P, \omega_o) = L_e(P, \omega_o) + \int_{S}f_s(P, \omega_i,
+    \omega_o) L_i(P, \omega_i) \vert cos\theta_i\vert d\omega_i
+  \end{align}
 
-    Rendering equation with solid angle:
-  \begin{align}
-    L_o(P, \omega_o) = L_o(P, \omega_o) + \int_{\Omega}f_s(P, \omega_i,
-    \omega_o) L_i(P, \omega_i) cos\Theta_i d\omega_i
-  \end{align}
-    To get from solid angle to path, consider:
+  Where:
+  	 $Lo(P,\:\omega_o)$ is the radiance at point $P$ in space towards direction $\omega_o$.
+
+  	 $L_e(P,\:\omega_o)$ is the emittance at point $P$ towards $\omega_o$ (such as a lamp).
+
+  	 $S$ is the unit sphere in which the angle $\omega_i$ is integrated over in respects to the normal at $P$.
+Also, this means, implicitly, that the equation is in respects to solid angle; the angle at which a primitive $A$ subtends $B$.
+
+  	 $f_s$ is the Bidirectional Scattering Distribution Function.
+This will be detailed in the BSDF chapter, suffice it to say, it describes how photons are scattered (or absorbed) from point P in space from direction $\omega_i$ towards $\omega_o$ -- either through reflection (BRDF), or transmittance (BTDF).
+In volumetric path tracing, there is also subsurface scattering.
+
+  	 $L_i(P, \omega_i)$ is the rendering equation at P from $\omega_i$
+
+  	 And $cos\theta_i$ is the normal attenuation.
+
+  Conceptually, this is the behavior of photons bouncing off several surfaces of a scene.
+While it's a very simple model, it's impossible to solve analytically for anything but the most simple of scenes, mostly due to, for example, the radiance at $P0$ relies on $P1$, and $P1$ relies on $P0$.
+
+  Another thing to remember is that while it is shown to be integrated over a sphere, it is almost always, and will be for the rest of this document, integrated over hemisphere $\Omega$ using either $f_r$ or $f_s$.
+
+  Unfortunately, this formula does not work very well in regards to bidirectional path tracing.
+The reason is because BDPT requires to work in terms of surfaces rather than solid angle, ei, $v_0 \rightarrow v_1$.
+So now the rendering equation must be transformed from solid angle to surfaces:
   \begin{align}

### 2017-11-27 testdocs/_region_.tex
doicumentation
whatever

--- 2017-11-24 documentation/_region_.tex
@@
-\message{ !name(paper.tex)}\documentclass{article}
+\message{ !name(test.tex)}\documentclass{article}
 \usepackage{amsmath}
@@
 
-\message{ !name(paper.tex) !offset(-3) }
+\message{ !name(test.tex) !offset(-3) }
 
@@
   \[\frac{(f(x+h) - f(x))}{h}\]
-  \begin{align*}
-    S(P, R) &= P - R^2   & \text{SDF of a sphere Radius  }\\
-    S(P, N) &= P \cdot N  & \text{SDF of a plane at Normal}\\
-    S(P, H, R) &= <|P_{x, z}| - R, \: P_{y}> - H & \text{SDF of a torus with
-                                                  Radius and Hole}\\
-  \end{align*}
-  \section{OpenCL}
-  \section{Light Transport Introduction}
-  \section{Bidirectional Path Tracing}
-    The main foundation of the mathematics and theory have come from many sources, from Veach's PhD thesis to PBR 3rd ed.
-and various open source path tracing kernels such as NanoRT, Embree, etc.
-However, the implementation differs from these in that the algorithm is optimized for the GPU.
-This section is about transforming the algorithms and mathematical models from Veach's thesis into something easily and efficiently implementable the GPU.
-
-    Calculating probability distribution functions is a very vital component in
-    Monte Carlo Sampling. The PDF states the probability of a random variable to
-    take the form of a given input. Its usefulness comes when contributions from
-    many Monte Carlo samples have a different weight. That is, one sample may
-    have a higher probability distribution, and thus a higher contribution to
-    the path. That's the strategy behind Monte Carlo;
-    \begin{align}
-      \int f(x) \approx \frac{1}{N}\Sigma_0^N\frac{f(x)}{p(x)}
-    \end{align}
-    The PDF applies to everything from path contributions, solid angles of a
-    BRDF, uniformly sampling a sphere, etc. I had a difficult time understanding
-    these, and more importantly, how to calculate them for a given formula. The
-    biggest conceptual barrier is that, for example, given a solid angle in a
-    hemisphere $\Omega$, the probability for a random variable X to equal
-    $\omega$ is 0. But this isn't what's happening here, this is all about the
-    probability that X could potentially equal $\omega$ when compared to other values
-    in $\Omega$. It's all about comparing the probabilities of two potential
-    $\omega$s with each other.
-
-    I still have a hard time conceptualizing these things in my head, but the
-    good thing is, is that calculating these are relatively easy. There's a lot
-    that could be talked about, but the important thing is that I've found a
-    reliable three-step process. The first is to
-    find the cumulative distribution function $CDF = Pr(X \leq x)$. Where $Pr$
-    is the probability, $X$ is a uniformly random variable, and $x$ is a known
-    constant. In other words, this gives the probability that, for a formula
-    $P$, where $X \in \Omega$, and $\Omega$ is the set of all possible values
-    for $CDF$,
-    that $X \leq x$. The second step is to take the derivative of this, so for
-    PDF $p$
-    \begin{align}
-      p(x) = \frac{dCDF(x)}{dx}
-    \end{align}
-    And the final step is to verify that the PDF is correct by integrating it
-    over $\Omega$, this final value should equal 1. This makes sense intuitively
-    -- it is a probability after all, so the sum of the probabilities of every
-    possible input must be 1.
-    For an example, how I calculated the PDF of an area light source
-      $L^0_e(v) = L_o(P, \omega_o)$ 
-    is quite simple. Firstly,the area light is defined as a sphere, and thus our
-    set of all possible values is correlated to surface area:
-    $\Omega = [0,\:2.0f*\tau*R^2]$.
-    Thus for a given constant $x$, and a uniformly random $X$,
-    \begin{align}
-      CDF(x) = Pr(X \leq x) = \frac{x}{2.0f*\tau*R^2}\\
-      p(x) = \frac{dCDF(x)}{dx} = \frac{1.0f}{2.0f*\tau*R^2}\\
-      \int_0^{\Omega}p(x)dx = 1.0f
-    \end{align}
-    In code
-    \begin{lstlisting}
-      float Light_PDF ( float3 P, Emitter* m ) {
-        float rad_sqr = m->radius*m->radius;
-        return 1.0f/(2.0f*TAU*rad_sqr);
-      }
-    \end{lstlisting}
-    The more interesting aspect, in this example, is actually generating the
-    uniformly distributed samples of a sphere. The naive method gives incorrect
-    distributions.TODO {page 774}
-
   The most obvious way to calculate this is the six-point gradient; where for
@@
 hand allows a cheap normal computation of primitives such as a spheres.
+  \section{OpenCL}
+	OpenCL, and GPU programming in general, is where I am least knowledgeable or interested.
+Well, OpenCL is easy to work with, and I wrote my own OpenCL module to handle OCL in a rather convenient method, for example, the DTOADQ kernel call looks as such:
+	\begin{lstlisting}
+	shared_info.image.Lock(); // Acquire OCL Mem from OGL image object
+	ocl.Run(
+	  ocl.CLStoreMem(shared_info.rw_image),
+	  shared_info.image.RWrite(),
+	  ocl.CLStoreMem(shared_info.image_metadata),
+	  shared_info.camera,
+	  shared_info.timer,
+	  ocl.CLPredefinedMem(imgs),
+	  ocl.CLStoreMem(shared_info.ocl_material),
+	  debug_values,
+	  ocl.CLStoreMem(shared_info.rng_states),
+	  // kernel size
+	  shared_info.image.x, shared_info.image.y
+	);
+	shared_info.image.Unlock();
 
-    Rendering equation with solid angle:
+	\end{lstlisting}
+
+	however the problem comes with dealing with the intrinsics of GPU; optimizing code for the GPU and knowing how memory and threads/workers are handled.
+
+  	One benefit of OpenCL is its ease to debug in comparison to OpenGL.
+One immediate benefit is printf, which surprisingly even works on the GPU.
+To make printf easily readable, I only use print on the center pixel of the image.
+Even so, due to the complexity of light transport, the vast majority of time spent on this project went directly towards debugging the DTOADQ kernel.
+
+  	OpenCL also makes it relatively easy to send and receive data between the host and GPU.
+In comparison to OpenGL, this is a big win, however, with OpenGL4.3's Shader Storage Buffer Object, this is now no longer a selling point.
+
+  	OpenCL also has huge drawbacks.
+The biggest problem is getting the results to be rendered in OpenGL.
+TODO.
+Also, there are minor annoyances, like the lack of a matrix data type, no function overloads, long compile times, or variadic functions.
+There are also annoyances with GPU programming in general.
+No dedicated random function.
+Though, possibly, long compile times is my fault for using a mega-kernel instead of splitting it into many small kernels.
+
+  	It seems that the GPU is not for me or DTOADQ and something I do not enjoy to use nor am good at.
+However, for this project, I had to deal with it.
+To overcome the lack of a random function, of which high quality uniform random generation is necessary for Monte Carlo simulations, I used the MWC64X algorithm.
+While there are better algorithms, they fail on branches which makes it useless for DTOADQ.
+It seemed to work well enough.
+The downside is that each worker/pixel needs its own random state, preserved across multiple kernel calls.
+There are indefinitely better solutions.
+  \section{Light Transport Introduction}
+  \section{Bidirectional Path Tracing}
+    The main foundation of the mathematics and theory have come from many sources, from Veach's PhD thesis to PBR 3rd ed.
+and various open source path tracing kernels such as NanoRT, Embree, etc.
+However, the implementation differs from these in that the algorithm is optimized for the GPU.
+This section is about transforming the algorithms and mathematical models from Veach's thesis into something easily and efficiently implementable the GPU.
+
+    Calculating probability distribution functions is a very vital component in
+    Monte Carlo Sampling. The PDF states the probability of a random variable to
+    take the form of a given input. Its usefulness comes when a Monte Carlo sample has a smaller or higher total contribution compared to other samples. That is, one sample may
+    have a higher probability distribution, and thus a higher contribution to
+    the path. That's the strategy behind Monte Carlo;
     \begin{align}
-    L_o(P, \omega_o) = L_o(P, \omega_o) + \int_{\Omega}f_s(P, \omega_i,
-    \omega_o) L_i(P, \omega_i) cos\Theta_i d\omega_i
+      \int f(x) \approx \frac{1}{N}\Sigma_0^N\frac{f(x)}{p(x)}
     \end{align}
-    To get from solid angle to path, consider:
+    The PDF applies to everything from path contributions, solid angles of a
+    BRDF, uniformly sampling a sphere, etc.
+
+    The
+    biggest conceptual barrier is that, for example, given a solid angle in a hemisphere $\Omega$, the probability for a random uniform $X$ to equal $\omega$ is 0.
+That's not what the PDF calculates, instead, it's the
+    probability that a random uniform $X$ could potentially equal $\omega$ when compared to other $\omega\:$s.
+
+    The good thing is that calculating these are relatively easy. The first step is to
+    find the cumulative distribution function $CDF = Pr(X \leq x)$. Where $Pr$
+    is the probability, $X$ is a uniformly random variable, and $x$ is a known
+    constant. This gives the probability that, for a formula
+    $P$, where $X \in \Omega$, and $\Omega$ is the set of all possible values
+    for $CDF$,
+    that $X \leq x$. Then the PDF $p$ is defined by the derivative of CDF. To verify that the PDF is correct, integrate it
+    over $\Omega$, this final value should equal 1. This makes sense intuitively
+    -- it is a probability after all, so the sum of the probabilities of every
+    possible input must be 1.
+    For an example, how I calculated the PDF of an area light source
+      $L^0_e(v) = L_o(P,\:\omega_o)$ 
+    is quite simple. Firstly,the area light is defined as a sphere, and thus our
+    set of all possible values is correlated to surface area:
+    $\Omega = [0,\:2.0f*\tau*R^2]$. Thus,
+    \begin{align}
+      CDF(x) = Pr(X \leq x) = \frac{x}{2.0*\tau*R^2}\\
+      p(x) = \frac{dCDF(x)}{dx} = \frac{1.0}{2.0*\tau*R^2}\\
+      \int_0^{\Omega}p(x)dx = 1.0
+    \end{align}
+    In code
+    \begin{lstlisting}
+      float Light_PDF ( Emitter* m ) {
+        return 1.0f/(2.0f*TAU*SQR(m->radius));
+      }
+    \end{lstlisting}
+
+  The rendering equation is:
+  \begin{align}
+    L_o(P, \omega_o) = L_e(P, \omega_o) + \int_{S}f_s(P, \omega_i,
+    \omega_o) L_i(P, \omega_i) \vert cos\theta_i\vert d\omega_i
+  \end{align}
+
+  Where:
+  	 $Lo(P,\:\omega_o)$ is the radiance at point $P$ in space towards direction $\omega_o$.
+
+  	 $L_e(P,\:\omega_o)$ is the emittance at point $P$ towards $\omega_o$ (such as a lamp).
+
+  	 $S$ is the unit sphere in which the angle $\omega_i$ is integrated over in respects to the normal at $P$.
+Also, this means, implicitly, that the equation is in respects to solid angle; the angle at which a primitive $A$ subtends $B$.
+
+  	 $f_s$ is the Bidirectional Scattering Distribution Function.
+This will be detailed in the BSDF chapter, suffice it to say, it describes how photons are scattered (or absorbed) from point P in space from direction $\omega_i$ towards $\omega_o$ -- either through reflection (BRDF), or transmittance (BTDF).
+In volumetric path tracing, there is also subsurface scattering.
+
+  	 $L_i(P, \omega_i)$ is the rendering equation at P from $\omega_i$
+
+  	 And $cos\theta_i$ is the normal attenuation.
+
+  Conceptually, this is the behavior of photons bouncing off several surfaces of a scene.
+While it's a very simple model, it's impossible to solve analytically for anything but the most simple of scenes, mostly due to, for example, the radiance at $P0$ relies on $P1$, and $P1$ relies on $P0$.
+
+  Another thing to remember is that while it is shown to be integrated over a sphere, it is almost always, and will be for the rest of this document, integrated over hemisphere $\Omega$ using either $f_r$ or $f_s$.
+
+  Unfortunately, this formula does not work very well in regards to bidirectional path tracing.
+The reason is because BDPT requires to work in terms of surfaces rather than solid angle, ei, $v_0 \rightarrow v_1$.
+So now the rendering equation must be transformed from solid angle to surfaces:
   \begin{align}
@@
   \end{align*}
+  [9]
+  Samuel D. Butler, Experimental and Theoretical Basis for a closed-Form Spectral BRDF model
+
+  [Henyey]
+   Diffuse Radation in the Galaxy, 1942. L, G. Henyey and J. L. Greenstein. 
 \end{document}
-
-\message{ !name(paper.tex) !offset(-459) }
+\message{ !name(test.tex) !offset(-562) }

### 2017-11-27 testdocs/test.tex
whatever

@@
 \usepackage[margin=1.15in]{geometry}
+\author{aodq}
 
@@
 While it's a very simple model, it's impossible to solve analytically for anything but the most simple of scenes, mostly due to, for example, the radiance at $P0$ relies on $P1$, and $P1$ relies on $P0$.
+Another thing to remember is that while it is shown to be integrated over a sphere, it is usually shown to be integrated over hemisphere $\Omega$ using either $f_r$ or $f_s$.
 
-  Another thing to remember is that while it is shown to be integrated over a sphere, it is almost always, and will be for the rest of this document, integrated over hemisphere $\Omega$ using either $f_r$ or $f_s$.
+  To compute this unbiasedly, it is required to use PDFs as described before to approximate the integral using Monte Carlo simulation.
+The technique is to generate $N$ uniformly random samples, weight them by their probability distribution, and then average the results:
 
-  Unfortunately, this formula does not work very well in regards to bidirectional path tracing.
-The reason is because BDPT requires to work in terms of surfaces rather than solid angle, ei, $v_0 \rightarrow v_1$.
-So now the rendering equation must be transformed from solid angle to surfaces:
+  \begin{align}
+   \int{f(x)} \: \approx \frac{1}{N}\Sigma_0^N{\frac{f(x)}{p(x)}}
+  \end{align}
+
+  With standard "backward" path tracing, a photon would be emitted from a light source and scatter in the scene until it hit an 'eye' or 'camera'.
+While this is what happens in reality, in computer graphics, most of the samples never end up hitting the camera/eye, so it is much more efficient to instead start from the eye, and end at an emitter/light source.
+In this case, since we are integrating over solid angle, the probably distribution function would be related to sampling the BRDF.
+So for example, on a specular surface, the PDF is $1.0$ where $\omega_o = reflect(\omega_i, N)$ and $0.0$ otherwise.
+An obvious optimization is to only sample over the valid BSDF angles, which is known as importance sampling.
+In DTQ, there are three BSDF functions:
+
+  \begin{list}{}{}
+  \item BSDF Spectrum: The albedo at $(P,\:\omega_i,\:\omega_o)$
+  \item BSDF Sample: Samples a $\omega_o$ for a $(P,\:\omega_i)$
+  \item BSDF PDF: The probability distribution of $(P,\:\omega_i,\:\omega_o)$
+  \end{list}
+
+  There are many more optimizations to be made as well.
+For example, at each scattering event, an additional photon can be sent towards each emitter.
+This is what is known as next-event estimation.
+While this improves performance drastically, naively averaging each contribution is a problem, because each path has a different probability to their paths/contributions.
+The technique to sample them properly, multiple importance sampling, is related to BDPT, however it will be detailed later.
+The important thing to understand is that, in this sampling strategy, there is a new PDF introduced which relates to the probability of a path being generated, and this is critical to understand the concept of BDPT.
+
+  Now, the optimization that is of focus in DTQ is Bidirectional Path Tracing.
+Instead of starting from an eye or light, a 'photon' (path) is generated from both, allowed to scatter in the scene for a few paths (the length is determined by Russian Roulette described later), and connected.
+It's not exactly the end-points of the paths that are being connected, but each vertex generated is connected to each other.
+[SHOW EXAM{E}
+
+  ]This actually has an interesting correlation to next-event estimation, as it could be seen as 'extending' the amount of emitters that can be connected to.
+
+  Unfortunately, this formula does not work with BDPT, as it is necessary to transform the formula to work in terms of surface areas rather than solid angle, ei, $v_0 \rightarrow v_1$.
+The transformation looks like so:
   \begin{align}
@@
     $P_{+1}$ ($\ddot{v}$). More accurately, it's the transport between
-    three different surfaces of the scene. In terms of SDFs, the surface
-    is every point on an SDF map where $|map| \leq \epsilon$. There's no longer
-    any integration over a hemisphere with solid angle $\omega_i$, it's now over
-    every single surface in the scene. This will be explained in detail later.
-\\
-    In the case of $L_e$,
-    DTOADQ [at least, for now], only has support for area-light. Thus consider:
+    three different surfaces of the scene. There's a small change to be made in the case of $L_e$, as DTOADQ [at least, for now], only has support for area-light. Thus consider:
 
@@
 
-    The cos term, used for the projected solid angle in irradiance, no longer
-    applies when dealing with area. This and $d\omega_i$ now become a
-    generalized geometry term, G, that converts a PDF with respect to solid area
-    using Jacobian mapping, which involves the inverse squared distance, and the
-    cosine angle between the geometric normals at $v$ and $\dot{v}$:
+    Instead of working with solid angles, we now work with a geometry term, G, that converts a PDF with respect to solid angle to surface area, which involves the inverse squared distance, and the cosine angle between the geometric normals at $v$ and $\dot{v}$:
 
@@
     where $s$ is the light-path length, and $t$ is the eye-path length. In
-    DTOADQ, they are limited to $s+t \leq 8$[TODO]. Finally, to transform
-    $\bar{v}$ into two seperate paths, $\frac{f_j(\bar{v}_{s,
+    DTOADQ, they are limited to $s+t \leq DEPTH$, a kernel-defined macro. Finally, to transform
+    $\bar{v}$ into two seperate paths, so that an eye and light path may be generated, $\frac{f_j(\bar{v}_{s,
         t})}{p(\bar{v}_{s, t})}$ is split into two with a connection edge $c$

### 2017-11-29 testdocs/_region_.tex
doicumentation

@@
   \section{Signed Distance Fields}
-  Signed Distance Fields (SDFs) are a mathematical tool to represent a 3D primitive by describing its surface via a distance field.
-Given an arbitrary point in space, the SDF of a primitive is the distance to its nearest point/surface.
-For example, the SDF of a sphere is $\Vert\:O\:\Vert - R$.
-With SDFs, it's possible to march a ray along the scene until it hits the surface of an SDF within a given bound, using the returned SDF distance as the distance to march.
-Some example SDFs are shown below:    \begin{align*}
+  Signed Distance Fields (SDFs) are a mathematical tool to represent a 3D
+primitive by describing its surface via a distance field. Given an arbitrary
+point in space, the SDF of a primitive is the distance to its nearest
+point/surface. For example, the SDF of a sphere is $\Vert\:O\:\Vert - R$. With
+SDFs, it's possible to march a ray along the scene until it hits the surface of
+an SDF within a given bound, using the returned SDF distance as the distance to
+march. Some example SDFs are shown below: \begin{align*}
   S(P, R) &= \Vert\:P\:\Vert - R & \text{Sphere with Radius}\\
@@
 
-  The biggest issue with SDFs is Lipschitz continuity.
-The SDFs themselves have a distance gradient, that is, their bound.
-The bounds of a SDF is expected to be at most 1 in order to remain Lipschitz continuous, as otherwise the SDF will appear deformed as shown below.
+  The biggest issue with SDFs is Lipschitz continuity. The SDFs themselves have
+a distance gradient, that is, their bound. The bounds of a SDF is expected to be
+at most 1 in order to remain Lipschitz continuous, as otherwise the SDF will
+appear deformed as shown below.
 
@@
 
-  This is a larger problem in Light Transport than it is in raytracing because small discontinuities can impact the lighting of an entire scene.
-And to add to that, most interesting operations (such as the meta balls shown in the image above) will violate Lipschitz.
-The easiest way to fix this is to multiply the distance field by a constant to stay below 1, but this is less of a fix and more of a hack.
+  This is a larger problem in Light Transport than it is in raytracing because
+small discontinuities can impact the lighting of an entire scene. And to add to
+that, most interesting operations (such as the meta balls shown in the image
+above) will violate Lipschitz. The easiest way to fix this is to multiply the
+distance field by a constant to stay below 1, but this is less of a fix and more
+of a hack.
 
-  In order to compute a normal of an SDF, you must perform a gradient approximation: the central distance onf the SDF at
-  P. For example, the difference quotient:
+  In order to compute a normal of an SDF, you must perform a gradient
+approximation: the central distance onf the SDF at P. For example, the
+difference quotient:
   \[\frac{(f(x+h) - f(x))}{h}\]
@@
   \section{OpenCL}
-	OpenCL, and GPU programming in general, is where I am least knowledgeable or interested.
-Well, OpenCL is easy to work with, and I wrote my own OpenCL module to handle OCL in a rather convenient method, for example, the DTOADQ kernel call looks as such:
+	OpenCL, and GPU programming in general, is where I am least knowledgeable or
+interested. Well, OpenCL is easy to work with, and I wrote my own OpenCL module
+to handle OCL in a rather convenient method, for example, the DTOADQ kernel call
+looks as such:
 	\begin{lstlisting}
@@
 
-	however the problem comes with dealing with the intrinsics of GPU; optimizing code for the GPU and knowing how memory and threads/workers are handled.
+	however the problem comes with dealing with the intrinsics of GPU; optimizing
+code for the GPU and knowing how memory and threads/workers are handled.
 
-  	One benefit of OpenCL is its ease to debug in comparison to OpenGL.
-One immediate benefit is printf, which surprisingly even works on the GPU.
-To make printf easily readable, I only use print on the center pixel of the image.
-Even so, due to the complexity of light transport, the vast majority of time spent on this project went directly towards debugging the DTOADQ kernel.
+  	One benefit of OpenCL is its ease to debug in comparison to OpenGL. One
+immediate benefit is printf, which surprisingly even works on the GPU. To make
+printf easily readable, I only use print on the center pixel of the image. Even
+so, due to the complexity of light transport, the vast majority of time spent on
+this project went directly towards debugging the DTOADQ kernel.
 
-  	OpenCL also makes it relatively easy to send and receive data between the host and GPU.
-In comparison to OpenGL, this is a big win, however, with OpenGL4.3's Shader Storage Buffer Object, this is now no longer a selling point.
+  	OpenCL also makes it relatively easy to send and receive data between the
+host and GPU. In comparison to OpenGL, this is a big win, however, with
+OpenGL4.3's Shader Storage Buffer Object, this is now no longer a selling point.
 
-  	OpenCL also has huge drawbacks.
-The biggest problem is getting the results to be rendered in OpenGL.
-TODO.
-Also, there are minor annoyances, like the lack of a matrix data type, no function overloads, long compile times, or variadic functions.
-There are also annoyances with GPU programming in general.
-No dedicated random function.
-Though, possibly, long compile times is my fault for using a mega-kernel instead of splitting it into many small kernels.
+  	OpenCL also has huge drawbacks. The biggest problem is getting the results
+to be rendered in OpenGL. TODO. Also, there are minor annoyances, like the lack
+of a matrix data type, no function overloads, long compile times, or variadic
+functions. There are also annoyances with GPU programming in general. No
+dedicated random function. Though, possibly, long compile times is my fault for
+using a mega-kernel instead of splitting it into many small kernels.
 
-  	It seems that the GPU is not for me or DTOADQ and something I do not enjoy to use nor am good at.
-However, for this project, I had to deal with it.
-To overcome the lack of a random function, of which high quality uniform random generation is necessary for Monte Carlo simulations, I used the MWC64X algorithm.
-While there are better algorithms, they fail on branches which makes it useless for DTOADQ.
-It seemed to work well enough.
-The downside is that each worker/pixel needs its own random state, preserved across multiple kernel calls.
+  	It seems that the GPU is not for me or DTOADQ and something I do not enjoy
+to use nor am good at. However, for this project, I had to deal with it. To
+overcome the lack of a random function, of which high quality uniform random
+generation is necessary for Monte Carlo simulations, I used the MWC64X
+algorithm. While there are better algorithms, they fail on branches which makes
+it useless for DTOADQ. It seemed to work well enough. The downside is that each
+worker/pixel needs its own random state, preserved across multiple kernel calls.
 There are indefinitely better solutions.
@@
   \section{Bidirectional Path Tracing}
-    The main foundation of the mathematics and theory have come from many sources, from Veach's PhD thesis to PBR 3rd ed.
-and various open source path tracing kernels such as NanoRT, Embree, etc.
-However, the implementation differs from these in that the algorithm is optimized for the GPU.
-This section is about transforming the algorithms and mathematical models from Veach's thesis into something easily and efficiently implementable the GPU.
+    The main foundation of the mathematics and theory have come from many
+sources, such as Veach's thesis to PBR 3rd ed. and various open source path
+tracing kernels such as NanoRT, Embree, etc. However, the implementation differs
+from these in that the algorithm is optimized for the GPU. This section is about
+transforming the algorithms and mathematical models into something easily and
+efficiently implementable the GPU, however I will work from the ground up in the
+theory and mathematics behind light transport.
 
@@
 Monte Carlo Sampling. The PDF states the probability of a random variable to
-    take the form of a given input. Its usefulness comes when a Monte Carlo sample has a smaller or higher total contribution compared to other samples. That is, one sample may
-    have a higher probability distribution, and thus a higher contribution to
-    the path. That's the strategy behind Monte Carlo;
+take the form of a given input. Its usefulness comes when a Monte Carlo
+sample has a smaller or higher total contribution compared to other samples.
+That is, one sample may have a higher probability distribution, and thus a
+higher contribution to the path. That's the strategy behind Monte Carlo;
     \begin{align}
@@
 
-    The
-    biggest conceptual barrier is that, for example, given a solid angle in a hemisphere $\Omega$, the probability for a random uniform $X$ to equal $\omega$ is 0.
-That's not what the PDF calculates, instead, it's the
-    probability that a random uniform $X$ could potentially equal $\omega$ when compared to other $\omega\:$s.
+    The biggest conceptual barrier is that, for example, given a solid angle in
+a hemisphere $\Omega$, the probability for a random uniform $X$ to equal
+$\omega$ is 0. That's not what the PDF calculates, instead, it's the probability
+that a random uniform $X$ could potentially equal $\omega$ when compared to
+other $\omega\:$s.
 
-    The good thing is that calculating these are relatively easy. The first step is to
-    find the cumulative distribution function $CDF = Pr(X \leq x)$. Where $Pr$
+    The good thing is that calculating these are relatively easy. The first step
+is to find the cumulative distribution function $CDF = Pr(X \leq x)$. Where $Pr$
 is the probability, $X$ is a uniformly random variable, and $x$ is a known
-    constant. This gives the probability that, for a formula
-    $P$, where $X \in \Omega$, and $\Omega$ is the set of all possible values
-    for $CDF$,
-    that $X \leq x$. Then the PDF $p$ is defined by the derivative of CDF. To verify that the PDF is correct, integrate it
-    over $\Omega$, this final value should equal 1. This makes sense intuitively
-    -- it is a probability after all, so the sum of the probabilities of every
-    possible input must be 1.
-    For an example, how I calculated the PDF of an area light source
+constant. This gives the probability that, for a formula $P$, where $X \in
+\Omega$, and $\Omega$ is the set of all possible values for $CDF$, that $X \leq
+x$. Then the PDF $p$ is defined by the derivative of CDF. To verify that the PDF
+is correct, integrate it over $\Omega$, this final value should equal 1. This
+makes sense intuitively -- it is a probability after all, so the sum of the
+probabilities of every possible input must be 1. For an example, how I
+calculated the PDF of an area light source
       $L^0_e(v) = L_o(P,\:\omega_o)$ 
@@
 
-  The rendering equation is:
+  Now, the importance of probability distribution functions comes when
+performing Monte Carlo approximation on an integral. Take, for instance, the
+most well-known form of the rendering equation:
   \begin{align}
@@
   Where:
-  	 $Lo(P,\:\omega_o)$ is the radiance at point $P$ in space towards direction $\omega_o$.
+  \begin{list}{}{}
+  	 \item $Lo(P,\:\omega_o)$ is the radiance at point $P$ in space towards direction $\omega_o$.
 
-  	 $L_e(P,\:\omega_o)$ is the emittance at point $P$ towards $\omega_o$ (such as a lamp).
+  	 \item $L_e(P,\:\omega_o)$ is the emittance at point $P$ towards $\omega_o$
+      (such as a lamp).
 
-  	 $S$ is the unit sphere in which the angle $\omega_i$ is integrated over in respects to the normal at $P$.
-Also, this means, implicitly, that the equation is in respects to solid angle; the angle at which a primitive $A$ subtends $B$.
+  	 \item $S$ is the unit sphere in which the angle $\omega_i$ is integrated
+        over in respects to the normal at $P$. Also, this means, implicitly,
+        that the equation is in respects to solid angle; the angle at which a
+        primitive $A$ subtends $B$.
 
-  	 $f_s$ is the Bidirectional Scattering Distribution Function.
-This will be detailed in the BSDF chapter, suffice it to say, it describes how photons are scattered (or absorbed) from point P in space from direction $\omega_i$ towards $\omega_o$ -- either through reflection (BRDF), or transmittance (BTDF).
-In volumetric path tracing, there is also subsurface scattering.
+  	 \item $f_s$ is the Bidirectional Scattering Distribution Function. This
+        will be detailed in the BSDF chapter, suffice it to say, it describes
+        how photons are scattered (or absorbed) from point P in space from
+        direction $\omega_i$ towards $\omega_o$ -- either through reflection
+        (BRDF), or transmittance (BTDF). In volumetric path tracing, there is
+        also subsurface scattering, but this phenomena is only approximated in
+        DTQ.
 
-  	 $L_i(P, \omega_i)$ is the rendering equation at P from $\omega_i$
+  	 \item $L_i(P, \omega_i)$ is the rendering equation at P from $\omega_i$
 
-  	 And $cos\theta_i$ is the normal attenuation.
+  	 \item And $cos\theta_i$ is the normal attenuation.
+  \end{list}
+  Conceptually, this is the behavior of photons bouncing off several surfaces of
+a scene. While it's a very simple model, it's impossible to solve analytically
+for anything but the most simple of scenes, mostly due to, for example, the
+radiance at $P0$ relies on $P1$, and $P1$ relies on $P0$. Another thing to
+remember is that while it is shown to be integrated over a sphere, it is usually
+shown to be integrated over hemisphere $\Omega$ using either $f_r$ or $f_s$.
 
-  Conceptually, this is the behavior of photons bouncing off several surfaces of a scene.
-While it's a very simple model, it's impossible to solve analytically for anything but the most simple of scenes, mostly due to, for example, the radiance at $P0$ relies on $P1$, and $P1$ relies on $P0$.
+  To compute this unbiasedly, it is required to use PDFs as described before to
+approximate the integral using Monte Carlo simulation. The technique is to
+generate $N$ uniformly random samples, weight them by their probability
+distribution, and then average the results:
 
-  Another thing to remember is that while it is shown to be integrated over a sphere, it is almost always, and will be for the rest of this document, integrated over hemisphere $\Omega$ using either $f_r$ or $f_s$.
+  \begin{align}
+   \int{f(x)} \: \approx \frac{1}{N}\Sigma_0^N{\frac{f(x)}{p(x)}}
+  \end{align}
 
-  Unfortunately, this formula does not work very well in regards to bidirectional path tracing.
-The reason is because BDPT requires to work in terms of surfaces rather than solid angle, ei, $v_0 \rightarrow v_1$.
-So now the rendering equation must be transformed from solid angle to surfaces:
+  With standard "backward" path tracing, a photon would be emitted from a light
+source and scatter in the scene until it hit an 'eye' or 'camera'. While this is
+what happens in reality, in computer graphics, most of the samples never end up
+hitting the camera/eye, so it is much more efficient to instead start from the
+eye, and end at an emitter/light source. In this case, since we are integrating
+over solid angle, the probably distribution function would be related to
+sampling the BRDF. So for example, on a specular surface, the PDF is $1.0$ where
+$\omega_o = reflect(\omega_i, N)$ and $0.0$ otherwise. An obvious optimization
+is to only sample over the valid BSDF angles, which is known as importance
+sampling. In DTQ, there are three BSDF functions:
+
+  \begin{list}{}{}
+  \item BSDF Spectrum: The albedo at $(P,\:\omega_i,\:\omega_o)$
+  \item BSDF Sample: Samples a $\omega_o$ given $(P,\:\omega_i)$
+  \item BSDF PDF: The probability distribution of $(P,\:\omega_i,\:\omega_o)$
+  \end{list}
+
+  There are many more optimizations to be made as well. For example, at each
+scattering event, an additional photon can be sent towards each emitter. This is
+what is known as next-event estimation. While this improves performance
+drastically, naively averaging each contribution is incorrect, because each path
+has a different probability to their paths/contributions. The technique to
+sample them properly, multiple importance sampling, is related to BDPT, however
+it will be detailed . The important thing to understand is that, in this
+sampling strategy, there is a new PDF introduced which relates to the
+probability of a path being generated, and this is critical to understand the
+concept of BDPT.
+
+  Now, the optimization that is of focus in DTQ is Bidirectional Path Tracing.
+Instead of starting from an eye or light, a 'photon' (path) is generated from
+both, allowed to scatter in the scene for a few paths (the length is determined
+by Russian Roulette), and connected. It's not exactly the end-points of the
+paths that are being connected, but each vertex generated is connected to each
+other. [SHOW EXAMPLE]. This actually has an interesting correlation to
+next-event estimation, as it could be seen as 'extending' the amount of emitters
+that can be connected to.
+
+  For BDPT, like a lot of different light transport equations, things go much
+more smoothly when the equation uses a different measure. In this case, the
+measure is over surface area rather than solid angles. This looks like $v_0
+\rightarrow v_1$. The transformation looks like so:
   \begin{align}
@@
     $P_{+1}$ ($\ddot{v}$). More accurately, it's the transport between
-    three different surfaces of the scene. In terms of SDFs, the surface
-    is every point on an SDF map where $|map| \leq \epsilon$. There's no longer
-    any integration over a hemisphere with solid angle $\omega_i$, it's now over
-    every single surface in the scene. This will be explained in detail later.
-\\
-    In the case of $L_e$,
-    DTOADQ [at least, for now], only has support for area-light. Thus consider:
+    three different surfaces of the scene. There's a small change to be made in
+the case of $L_e$, as DTOADQ [at least, for now], only has support for
+area-light. Thus consider:
 
@@
 
-    The cos term, used for the projected solid angle in irradiance, no longer
-    applies when dealing with area. This and $d\omega_i$ now become a
-    generalized geometry term, G, that converts a PDF with respect to solid area
-    using Jacobian mapping, which involves the inverse squared distance, and the
-    cosine angle between the geometric normals at $v$ and $\dot{v}$:
+    The $cos\theta_i$ is no longer applicable, instead there is now a geometry
+term, G, that converts a PDF with respect to solid angle to surface area, which
+involves the inverse squared distance, and the cosine angle between the
+geometric normals at $v$ and $\dot{v}$:
 
@@
   \end{align}
-    where $A$ is the area on $M$, and $M$ is the union of all scene surfaces.
-    This is known as the three-point form or the light transport equation. By
-    recursively expanding$^{[1]}$, so that the three point form is integrated
-    over every possible set of possible paths, this can be rewritten as the more
-    usable format:
+    where $A$ is the area-measurement function on $M$, and $M$ is the set of all
+points on the surface of a scene. That is, to say in regards to SDFs, $M$ is the
+set of all points $P$ where $|map(P)| < \epsilon$. The area-measurement function
+goes into measure theory, which while it is beyond my grasp currently to
+compute, conceptually I understand a bit of it. The important thing is that $A$
+is a function on the set of all subsets of $M$ that returns the measurement of
+each set, and as long as the surfaces of the scene are sampled uniformly, this
+never actually has to be computed. To back it up a little further, Veach
+presented this as a means to integrate over a measure rather than using solid
+angles. This proves that you can transform the rendering equation to different
+measurements. For example, instead of integrating over points on surfaces, one
+could integrate over an entire 3D space in order to perform volumetric path
+tracing. Transforming the rendering equation using measurements would be
+interesting in the future to look into.
+
+    Anyways, with the three-point form now described, by recursively expanding,
+so that the three point form is integrated over every possible set of possible
+paths, this can be rewritten as the more usable format:
   \begin{align} I_j = \int_\Omega f_j(\bar{v}) d\mu(\bar{v}) \end{align}
@@
     paths of any length, and $\mu$ is the area-product measure [the product of
-    all the expanded $d_A$]. Special care has to be taken with the definition of
+all the expanded $A$]. Special care has to be taken with the definition of
     path itself, as in the source code, there never exists a path $\bar{v}$,
-    only eye-path $\bar{z}$ and light-path . Right now, this is analytically
-    unsolvable. To compute this, we need to apply monte carlo sampling and limit
-    the maximum path length:
+    only eye-path $\bar{z}$ and light-path . Monte Carlo approximation, as shown
+above, is applied to the formula:
   \begin{align}
@@
     where $s$ is the light-path length, and $t$ is the eye-path length. In
-    DTOADQ, they are limited to $s+t \leq 8$[TODO]. Finally, to transform
-    $\bar{v}$ into two seperate paths, $\frac{f_j(\bar{v}_{s,
-        t})}{p(\bar{v}_{s, t})}$ is split into two with a connection edge $c$
+    DTOADQ, they are limited to $s+t \leq DEPTH$, a kernel-defined macro.
+Finally, to transform $\bar{v}$ into two seperate paths, so that an eye and
+light path may be generated, $\frac{f_j(\bar{v}_{s, t})}{p(\bar{v}_{s, t})}$ is
+split into two with a connection edge $c$
   \begin{align}
@@
   \end{align}
+
+  $\alpha_{s|t}^{L|E}$ is the final radiance of either the emitter or eye path,
+  and thus it is recursive in nature. That is, the value of $\alpha_i$ is
+  dependent on $\alpha_{i-1}$, unless $i = 0$ in which case is the radiance of
+  either the emitter or the eye path. This is shown below.
   \begin{align}
@@
   \end{align}
-    There is no case for $i = 0$ as either, the path doesn't exist, or if one
-    does exist, $y_0$ and $z_0$ do not contribute to the image. Veach (and many other
-    sources), leave this in, but it's unnecessarily confusing. A special case
-    needs to be handled, unfortunately, for $\alpha^E_0$ as it is being
-    generated on the fly while $\bar{z}$ is being constructed. The two options
-    is to unroll the first iteration of the construction loop, or just allow the
-    special case to exist. The latter is the better choice, as all kernels will
-    enter and exit the special case at the same time, and the GPU might just
-    unroll the loop anyways.
-    There may be cases where $s = 0$ ($t = 0$ is not possible without a
-    physical camera lens), which is equivalent to just forward path tracing.
-    There is also to consider, what happens on $s = 1$ and $t = 1$, %TODO
-    Well, in either case where both paths exist or only the camera path exists,
-    the connection strategy below describes how to connect the edges of the two
-    paths
+    For the case of $i = 0$, this would mean that the path does not exist. In
+other words, forward or backward path tracing, the latter of which isn't
+supported as camera lens are not defined in DTQ. Veach (and many other sources),
+leave this in, but it's unnecessarily confusing. For $i = 1$, it is either a
+point on the surface of a light, or the origin point on the 'delta eye'. The
+connection strategy below describes how to connect the edges of the two paths
   \begin{align}
@@
   \end{align}
-    Is the $s = 0, t > 1$
-    strategy worthwhile to compute? Those familiar with next event
-    estimation would know that there are situations in which this is beneficial.
-    Take, for example, a glossy surface with a small solid angle subtending to
-    the surface of a large light source. The $s=1$ sample strategy in this case
-    would only contribute when the $L0$ vertex is within that solid angle.
-    The effective strategy to sample in this case is $s = 0$.
-    In code, $\alpha_i^E\:c_{s,t}\:\alpha_0^L$ looks like (MIS not shown):
+    Is the $s = 0, t > 1$ strategy worthwhile to compute? Those familiar with
+next event estimation would know that there are situations in which this is
+beneficial. Take, for example, a glossy surface with a small solid angle
+subtending to the surface of a large light source. The $s=1$ sample strategy in
+this case would only contribute when the $L0$ vertex is within that solid angle.
+The effective strategy to sample in this case is $s = 0$. In code,
+$\alpha_i^E\:c_{s,t}\:\alpha_0^L$ looks like (MIS not shown):
     \begin{lstlisting}
@@
     Another question to consider is if the path should be able to reflect on the
-    light source. While under real circumstances, this is allowed, the
-    contribution is negligible and dismissed in DTOADQ. Also, there are no cases for $s = s, t = 0|1$ as
-    the lens does not have a physical existence in the SDF map.
-The $t = 0$ case, where a photon from the light source hits the lens directly, is rather pointless anyways, as a path generated by light-tracing would have to be able to contribute to any pixel, which is not feasible on a GPU.
-The case of $t = 1$, where the light path is connected to a position on the lens, is important for effects such as depth of field and aperture.
-The only problem is due to time constraints, this has not been implemented yet.
+light source. While under real circumstances, this is allowed, the contribution
+is negligible and dismissed in DTOADQ. Also, there are no cases for $s = s, t =
+0|1$ as the lens does not have a physical existence in the SDF map. The $t = 0$
+case, where a photon from the light source hits the lens directly, is rather
+pointless anyways, as a path generated by light-tracing would have to be able to
+contribute to any pixel, which is not feasible on a GPU. The case of $t = 1$,
+where the light path is connected to a position on the lens, is important for
+effects such as depth of field and aperture. The only problem is due to time
+constraints, this has not been implemented yet.
 
     One immediately obvious optimization that could be made with this model is
-    in regards to the visibility check in the geometric term, if you were to
-    expand $I_j$, and take the V term from outside the G function
+in regards to the visibility check in the geometric term, if you were to expand
+$I_j$, and take the V term from outside the G function
     [$G(v \rightarrow
     \dot{v}) V(v \rightarrow \dot{v})$],
-    it would be made obvious that all the visibility checks can be
-    cancelled out for the entire equation, except for the connection term $c$.
-    This is intuitive as the paths
-    $\bar{y}_{0} \ldots \bar{y}_s$ and $\bar{z}_0 \ldots \bar{z}_t$ had just been generated via raymarching.
-    Specifically, the definition of $V$ is
+    it would be made obvious that all the visibility checks can be cancelled out
+for the entire equation, except for the connection term $c$. This is intuitive
+as the paths $\bar{y}_{0} \ldots \bar{y}_s$ and $\bar{z}_0 \ldots \bar{z}_t$ had
+just been generated via raymarching. Specifically, the definition of $V$ is
   \begin{align}
@@
     Moving on, for $\alpha^{L|E}_i$, these are equivalent for both $L$ and $E$,
-    the problem is that the $\bar{y}$ path is generated in perspective of the
-    light source, so there has to be a bit of caution in terms of perspective in when generating the PDFs.
-
-     $p_A$ is
-    the PDF w.r.t. area, for the light path; we only need to take the PDF of the
-    solid angle of a sphere
+the problem is that the $\bar{y}$ path is generated in perspective of the light
+source, so there has to be a bit of caution in terms of perspective in when
+generating the PDFs. $p_A$ is the PDF w.r.t. area, for the light path; we only
+need to take the PDF of the solid angle of a sphere
   \begin{align}
@@
     where $R$ is the radius of the sphere, note $y_0$ is a point on the surface
-    of the sphere, and not its origin.
-    \\$p_A^E(x_0 \rightarrow x_1)$ doesn't exist, as there is no physical lens
-    on the scene. [TODO - this might cahnge later]. One of the benefits of this,
-    is that there are no special cases for pathes where $t = 0$, which is
-    impossible to handle on the GPU as the path could contribute to any
-    arbitrary pixel [directly hitting the lens].
+of the sphere, and not its origin. \\$p_A^E(x_0 \rightarrow x_1)$ doesn't exist,
+as there is no physical lens on the scene. [TODO - this might cahnge later]. One
+of the benefits of this, is that there are no special cases for pathes where $t
+= 0$, which is impossible to handle on the GPU as the path could contribute to
+any arbitrary pixel [directly hitting the lens].
   \begin{align}
@@
 only supports area lights, so $L^0_e$ is set a constant for the emitter's
-    spectrum value.
-    The $W_e^0$ is contribution from the eye source, that is, the lens. Since
-    the lens does not exist in the scene, as of now [TODO - this might change
-    later], there is no $W_e^0$. While this should be implemented in the future,
-    a hidden benefit of the current model is that there are no special cases for
-    paths where $t = 0$, which is impossible to handle on the GPU as any such
+spectrum value. The $W_e^0$ is contribution from the eye source, that is, the
+lens. Since the lens does not exist in the scene, as of now [TODO - this might
+change later], there is no $W_e^0$. While this should be implemented in the
+future, a hidden benefit of the current model is that there are no special cases
+for paths where $t = 0$, which is impossible to handle on the GPU as any such
 path could contribute to any arbitrary pixel [the light path is, after all,
@@
   \end{align}
-  The amount that each path contributes to F differ greatly, so naively
-  weighing each contribution uniformly by something like 1.0f, or even, by
-  inverse path length, will producy very noisy images, and would not be
-  worth the additional computation time that BDPT requires, as shown below.
-  	 To calculate the PDF of 
-  In order to get good
-  samples from the contributions, they must be combined in an optimal manner
-  using Multiple Importance Sampling: each contribution is weighted by the PDF
-  of the entire path itself:
+  The amount that each path contributes to F differ greatly, so naively weighing
+each contribution uniformly by something like 1.0f, or even, by inverse path
+length, will producy very noisy images, and would not be worth the additional
+computation time that BDPT requires, as shown below. To calculate the PDF of In
+order to get good samples from the contributions, they must be combined in an
+optimal manner using Multiple Importance Sampling: each contribution is weighted
+by the PDF of the entire path itself:
   \begin{align*}
@@
 strategies that could could have, in theory,$^{[1]}$ created the path. The
-    straight forward implementation of this has a lot of problems gone
-    unresolved in Veach's thesis; PBR 3rd Ed, chapter 16 pg 1014-1016 addresses
-    and describe these problems. Primarily, $P_i(\bar{x})$ will overflow
-    CLFloat (due to the distance in the Geometric term), and the straight
-    forward implementation is very long, hard to debug, and has a poor time
-    complexity. But the important part is the end result of their solution:
-
+straight forward implementation of this has a lot of problems gone unresolved in
+Veach's thesis; PBR 3rd Ed, chapter 16 pg 1014-1016 addresses and describe these
+problems. Primarily, $P_i(\bar{x})$ will overflow CLFloat (due to the distance
+in the Geometric term), and the straight forward implementation is very long,
+hard to debug, and has a poor time complexity. But the important part is the end
+result of their solution:
   \begin{align}
@@
   \end{align}
-
     However, DTOADQ can simplify a bit further; unlike above, there are two
-    paths for which to evaluate. Specifically, in the above term, for where
-    $i < s$, we have
-    $\frac{\overleftarrow{p}(\bar{x}_i)}{\overrightarrow{p}(\bar{x}_i)}$, 
+paths for which to evaluate. Specifically, in the above term, for where $i < s$,
+we have $\frac{\overleftarrow{p}(\bar{x}_i)}{\overrightarrow{p}(\bar{x}_i)}$,
 which is the light path (i is iterating towards light-length s). The eye
@@
 $\bar{x}_{i-1}$ is used). Another way of saying it, is $\bar{x} = \bar{y} ||
-    \bar{z}_{rev}$.
-    In DTOADQ's case, the weights are calculated in
-    the same direction relative to its originator; in this case the perspective
-    of the eye path.
+\bar{z}_{rev}$. In DTOADQ's case, the weights are calculated in the same
+direction relative to its originator; in this case the perspective of the eye
+path.
   \begin{align}
@@
     And then the expanded form of $\mathcal{W}_i(\bar{x})$ looks like
-
   \begin{align}
@@
   \end{align}
-
     And now the MIS looks like:
@@
     Thus, the final rendering equation for DTOADQ looks like such:
-
   \begin{align}
@@
   \section{DTOADQ BRDF}
-  DTOADQ supports a "universal" physically-based sampling and spectrum BRDF.
-The spectrum BRDF is a combination of many different modified BRDF functions into one.
-There is a microfacet, subsurface and diffusive component in the DTOADQ spectrum BRDF, which are all combined together, with many variables to tune these components.
-There is diffusive, glossy, specular and transmittive sampling in the BRDF importance sample, with variables to allow specific control over which strategies to choose and how they are combined together.
-Regardless of how the BRDF is sampled, they use the same spectrum BRDF.
+  DTOADQ supports a "universal" physically-based sampling and spectrum BRDF. The
+spectrum BRDF is a combination of many different modified BRDF functions into
+one. There is a microfacet, subsurface and diffusive component in the DTOADQ
+spectrum BRDF, which are all combined together, with many variables to tune
+these components. There is diffusive, glossy, specular and transmittive sampling
+in the BRDF importance sample, with variables to allow specific control over
+which strategies to choose and how they are combined together. Regardless of how
+the BRDF is sampled, they use the same spectrum BRDF.
 
-  In order to develop these BRDFs, I wrote an SDF model based off the mitsuba model and wrote an IBL path-traced shader for quick prototyping; shadertoy.com/view/XtSyDm .
-The main reason to not use DTOADQ was the lack of cubemap support, which would have taken too long to implement.
+  In order to develop these BRDFs, I wrote an SDF model based off the mitsuba
+model and wrote an IBL path-traced shader for quick prototyping;
+shadertoy.com/view/XtSyDm . The main reason to not use DTOADQ was the lack of
+cubemap support, which would have taken too long to implement.
 
   The Microfacet model consists of a fresnel, distribution and geometric term.
-It's important to note that these have all been modified to use the half-vector $H = \Vert\:L+V\:\Vert$.
-This gives better visuals and allows for a specific energy conservation model.
-Anyways, the fresnel term approximates the fresnel factor, it's based off both Schlick's approximation and Disney's diffusive schlick fresnel, with modifications to use a $metallic$ term as well as an albedo;
+It's important to note that these have all been modified to use the half-vector
+$H = \Vert\:L+V\:\Vert$. This gives better visuals and allows for a specific
+energy conservation model. Anyways, the fresnel term approximates the fresnel
+factor, it's based off both Schlick's approximation and Disney's diffusive
+schlick fresnel, with modifications to use a $metallic$ term as well as an
+albedo;
   \begin{lstlisting}
@@
 
-  The Geometric term is used to control how microfacets 'shadow' each other.
-The DTQ BRDF uses Heits' [2014] SmithGGXCorrelated modified to use a half vector, combined with an anisotropic term based off the GGX Anisotropic model:
+  The Geometric term is used to control how microfacets 'shadow' each other. The
+DTQ BRDF uses Heits' [2014] SmithGGXCorrelated modified to use a half vector,
+combined with an anisotropic term based off the GGX Anisotropic model:
   \begin{lstlisting}
@@
   \includegraphics[scale=0.20]{geometric}
-
-  And finally, the Distribution term controls how microfacets are spread around the normal.
-It's based off the Hyper Cauchy with a roughness and metallic (shape) parameters [9].
-The latter of which should most likely be its own parameter in the future.
-The more readable mathematical model is shown:
-
+  And finally, the Distribution term controls how microfacets are spread around
+the normal. It's based off the Hyper Cauchy with a roughness and metallic
+(shape) parameters [9]. The latter of which should most likely be its own
+parameter in the future. The more readable mathematical model is shown:
   \begin{align}
@@
 
-  These combine to form the microfacet under the half vector microfacet energy conservation model [Edwards et al. 2006].
+  These combine to form the microfacet under the half vector microfacet energy
+conservation model [Edwards et al. 2006].
   \[
@@
 
-  The diffusive component is a combination of both the subsurface and lambertian component.
-The lambertian is simply $\alpha^{0.5} * \frac{1}{\pi}$, and the subsurface component is based off the Henyey-Greenstein formula [Henyey], modified for half-vec:
+  The diffusive component is a combination of both the subsurface and lambertian
+component. The lambertian is simply $\alpha^{0.5} * \frac{1}{\pi}$, and the
+subsurface component is based off the Henyey-Greenstein formula [Henyey],
+modified for half-vec:
   \begin{align}
@@
   \end{align}
-  where m is metallic and r is roughness. This is then combined to the diffusive component of the BRDF using Fresnel factor like so:
+  where m is metallic and r is roughness. This is then combined to the diffusive
+component of the BRDF using Fresnel factor like so:
   \begin{lstlisting}
@@
   \section{Bibliography}
-  [1]
-  \begin{align*}
-    \Sigma_{k=1}^{\infty} \int_{M^{k+1}} L_e(v_0 \rightarrow v_1) G(v_0
-    \leftrightarrow v_1)
-    \Pi_{i=1}^{k-1} F_s(v_{i-1} \rightarrow v_i \rightarrow v_{i+1}) G(v_i
-    \leftrightarrow v_{i+1})\\
-    W_e^j(v_{k-1} \rightarrow v_k) dA(v_0) \ldots dA(v_k)
-  \end{align*}
   [9]
@@
 \end{document}
-\message{ !name(test.tex) !offset(-562) }
+\message{ !name(test.tex) !offset(-684) }

### 2017-11-29 testdocs/test.tex
doicumentation

@@
   Where:
-  	 $Lo(P,\:\omega_o)$ is the radiance at point $P$ in space towards direction $\omega_o$.
+  \begin{list}{}{}
+  	 \item $Lo(P,\:\omega_o)$ is the radiance at point $P$ in space towards direction $\omega_o$.
 
-  	 $L_e(P,\:\omega_o)$ is the emittance at point $P$ towards $\omega_o$ (such as a lamp).
+  	 \item $L_e(P,\:\omega_o)$ is the emittance at point $P$ towards $\omega_o$
+      (such as a lamp).
 
-  	 $S$ is the unit sphere in which the angle $\omega_i$ is integrated over in respects to the normal at $P$.
-Also, this means, implicitly, that the equation is in respects to solid angle; the angle at which a primitive $A$ subtends $B$.
+  	 \item $S$ is the unit sphere in which the angle $\omega_i$ is integrated
+        over in respects to the normal at $P$. Also, this means, implicitly,
+        that the equation is in respects to solid angle; the angle at which a
+        primitive $A$ subtends $B$.
 
-  	 $f_s$ is the Bidirectional Scattering Distribution Function.
-This will be detailed in the BSDF chapter, suffice it to say, it describes how photons are scattered (or absorbed) from point P in space from direction $\omega_i$ towards $\omega_o$ -- either through reflection (BRDF), or transmittance (BTDF).
-In volumetric path tracing, there is also subsurface scattering.
+  	 \item $f_s$ is the Bidirectional Scattering Distribution Function. This
+        will be detailed in the BSDF chapter, suffice it to say, it describes
+        how photons are scattered (or absorbed) from point P in space from
+        direction $\omega_i$ towards $\omega_o$ -- either through reflection
+        (BRDF), or transmittance (BTDF). In volumetric path tracing, there is
+        also subsurface scattering, but this phenomena is only approximated in
+        DTQ.
 
-  	 $L_i(P, \omega_i)$ is the rendering equation at P from $\omega_i$
+  	 \item $L_i(P, \omega_i)$ is the rendering equation at P from $\omega_i$
 
-  	 And $cos\theta_i$ is the normal attenuation.
+  	 \item And $cos\theta_i$ is the normal attenuation.
+  \end{list}
+  Conceptually, this is the behavior of photons bouncing off several surfaces of
+a scene. While it's a very simple model, it's impossible to solve analytically
+for anything but the most simple of scenes, mostly due to, for example, the
+radiance at $P0$ relies on $P1$, and $P1$ relies on $P0$. Another thing to
+remember is that while it is shown to be integrated over a sphere, it is usually
+shown to be integrated over hemisphere $\Omega$ using either $f_r$ or $f_s$.
 
-  Conceptually, this is the behavior of photons bouncing off several surfaces of a scene.
-While it's a very simple model, it's impossible to solve analytically for anything but the most simple of scenes, mostly due to, for example, the radiance at $P0$ relies on $P1$, and $P1$ relies on $P0$.
-Another thing to remember is that while it is shown to be integrated over a sphere, it is usually shown to be integrated over hemisphere $\Omega$ using either $f_r$ or $f_s$.
-
-  To compute this unbiasedly, it is required to use PDFs as described before to approximate the integral using Monte Carlo simulation.
-The technique is to generate $N$ uniformly random samples, weight them by their probability distribution, and then average the results:
+  To compute this unbiasedly, it is required to use PDFs as described before to
+approximate the integral using Monte Carlo simulation. The technique is to
+generate $N$ uniformly random samples, weight them by their probability
+distribution, and then average the results:
 
@@
 
-  With standard "backward" path tracing, a photon would be emitted from a light source and scatter in the scene until it hit an 'eye' or 'camera'.
-While this is what happens in reality, in computer graphics, most of the samples never end up hitting the camera/eye, so it is much more efficient to instead start from the eye, and end at an emitter/light source.
-In this case, since we are integrating over solid angle, the probably distribution function would be related to sampling the BRDF.
-So for example, on a specular surface, the PDF is $1.0$ where $\omega_o = reflect(\omega_i, N)$ and $0.0$ otherwise.
-An obvious optimization is to only sample over the valid BSDF angles, which is known as importance sampling.
-In DTQ, there are three BSDF functions:
+  With standard "backward" path tracing, a photon would be emitted from a light
+source and scatter in the scene until it hit an 'eye' or 'camera'. While this is
+what happens in reality, in computer graphics, most of the samples never end up
+hitting the camera/eye, so it is much more efficient to instead start from the
+eye, and end at an emitter/light source. In this case, since we are integrating
+over solid angle, the probably distribution function would be related to
+sampling the BRDF. So for example, on a specular surface, the PDF is $1.0$ where
+$\omega_o = reflect(\omega_i, N)$ and $0.0$ otherwise. An obvious optimization
+is to only sample over the valid BSDF angles, which is known as importance
+sampling. In DTQ, there are three BSDF functions:
 
@@
   \item BSDF Spectrum: The albedo at $(P,\:\omega_i,\:\omega_o)$
-  \item BSDF Sample: Samples a $\omega_o$ for a $(P,\:\omega_i)$
+  \item BSDF Sample: Samples a $\omega_o$ given $(P,\:\omega_i)$
   \item BSDF PDF: The probability distribution of $(P,\:\omega_i,\:\omega_o)$
@@
 
-  There are many more optimizations to be made as well.
-For example, at each scattering event, an additional photon can be sent towards each emitter.
-This is what is known as next-event estimation.
-While this improves performance drastically, naively averaging each contribution is a problem, because each path has a different probability to their paths/contributions.
-The technique to sample them properly, multiple importance sampling, is related to BDPT, however it will be detailed later.
-The important thing to understand is that, in this sampling strategy, there is a new PDF introduced which relates to the probability of a path being generated, and this is critical to understand the concept of BDPT.
+  There are many more optimizations to be made as well. For example, at each
+scattering event, an additional photon can be sent towards each emitter. This is
+what is known as next-event estimation. While this improves performance
+drastically, naively averaging each contribution is incorrect, because each path
+has a different probability to their paths/contributions. The technique to
+sample them properly, multiple importance sampling, is related to BDPT, however
+it will be detailed . The important thing to understand is that, in this
+sampling strategy, there is a new PDF introduced which relates to the
+probability of a path being generated, and this is critical to understand the
+concept of BDPT.
 
   Now, the optimization that is of focus in DTQ is Bidirectional Path Tracing.
-Instead of starting from an eye or light, a 'photon' (path) is generated from both, allowed to scatter in the scene for a few paths (the length is determined by Russian Roulette described later), and connected.
-It's not exactly the end-points of the paths that are being connected, but each vertex generated is connected to each other.
-[SHOW EXAM{E}
+Instead of starting from an eye or light, a 'photon' (path) is generated from
+both, allowed to scatter in the scene for a few paths (the length is determined
+by Russian Roulette), and connected. It's not exactly the end-points of the
+paths that are being connected, but each vertex generated is connected to each
+other. [SHOW EXAMPLE]. This actually has an interesting correlation to
+next-event estimation, as it could be seen as 'extending' the amount of emitters
+that can be connected to.
 
-  ]This actually has an interesting correlation to next-event estimation, as it could be seen as 'extending' the amount of emitters that can be connected to.
-
-  Unfortunately, this formula does not work with BDPT, as it is necessary to transform the formula to work in terms of surface areas rather than solid angle, ei, $v_0 \rightarrow v_1$.
-The transformation looks like so:
+  For BDPT, like a lot of different light transport equations, things go much
+more smoothly when the equation uses a different measure. In this case, the
+measure is over surface area rather than solid angles. This looks like $v_0
+\rightarrow v_1$. The transformation looks like so:
   \begin{align}
@@
     $P_{+1}$ ($\ddot{v}$). More accurately, it's the transport between
-    three different surfaces of the scene. There's a small change to be made in the case of $L_e$, as DTOADQ [at least, for now], only has support for area-light. Thus consider:
+    three different surfaces of the scene. There's a small change to be made in
+the case of $L_e$, as DTOADQ [at least, for now], only has support for
+area-light. Thus consider:
 
@@
 
-    Instead of working with solid angles, we now work with a geometry term, G, that converts a PDF with respect to solid angle to surface area, which involves the inverse squared distance, and the cosine angle between the geometric normals at $v$ and $\dot{v}$:
+    The $cos\theta_i$ is no longer applicable, instead there is now a geometry
+term, G, that converts a PDF with respect to solid angle to surface area, which
+involves the inverse squared distance, and the cosine angle between the
+geometric normals at $v$ and $\dot{v}$:
 
@@
     where $s$ is the light-path length, and $t$ is the eye-path length. In
-    DTOADQ, they are limited to $s+t \leq DEPTH$, a kernel-defined macro. Finally, to transform
-    $\bar{v}$ into two seperate paths, so that an eye and light path may be generated, $\frac{f_j(\bar{v}_{s,
-        t})}{p(\bar{v}_{s, t})}$ is split into two with a connection edge $c$
+    DTOADQ, they are limited to $s+t \leq DEPTH$, a kernel-defined macro.
+Finally, to transform $\bar{v}$ into two seperate paths, so that an eye and
+light path may be generated, $\frac{f_j(\bar{v}_{s, t})}{p(\bar{v}_{s, t})}$ is
+split into two with a connection edge $c$
   \begin{align}
@@
   \end{align}
+
+  $\alpha_{s|t}^{L|E}$ is the final radiance of either the emitter or eye path,
+  and thus it is recursive in nature. That is, the value of $\alpha_i$ is
+  dependent on $\alpha_{i-1}$, unless $i = 0$ in which case is the radiance of
+  either the emitter or the eye path. However, this can be simplified further,
+because of Helmhotz reciprocity, BRDFs return the same value if $wo$ and $wi$
+are switched. This is also the reason why forward and backward path tracing
+work.
   \begin{align}
@@
     \begin{cases}
-      \frac{L_e^0(y_0 \rightarrow y_1)}{p_A(y_1)} | \frac{W^0_e(z_0)}{p_A(z_0)},
-      &i = 1,\\
-      (\frac{f_s(y_{i+1} \rightarrow y_i \rightarrow y_{i-1})}{p_{\sigma}(y_i
-      \rightarrow y_{i-1})} | \frac{f_s(z_{i-1} \rightarrow z_{i-2} \rightarrow
-      z_{i-3}}{p_\sigma(z_{i-2} \rightarrow z_{i-1})}) \alpha^{L|E}_{i-1},
-      &i > 1
+      1.0 &i = 0,\\
+      \frac{
+      L_e^0(y_0 \rightarrow y_1)}{p_A(y_1)} | 1.0 &i = 1,\\
+      \frac{f_s(v_{i-1} \rightarrow v_i \rightarrow v_{i+1})}{p_{\sigma}(v_i
+      \rightarrow v_{i-1})} &i > 1
     \end{cases}
   \end{align}
-    There is no case for $i = 0$ as either, the path doesn't exist, or if one
-    does exist, $y_0$ and $z_0$ do not contribute to the image. Veach (and many other
-    sources), leave this in, but it's unnecessarily confusing. A special case
-    needs to be handled, unfortunately, for $\alpha^E_0$ as it is being
-    generated on the fly while $\bar{z}$ is being constructed. The two options
-    is to unroll the first iteration of the construction loop, or just allow the
-    special case to exist. The latter is the better choice, as all kernels will
-    enter and exit the special case at the same time, and the GPU might just
-    unroll the loop anyways.
-    There may be cases where $s = 0$ ($t = 0$ is not possible without a
-    physical camera lens), which is equivalent to just forward path tracing.
-    There is also to consider, what happens on $s = 1$ and $t = 1$, %TODO
-    Well, in either case where both paths exist or only the camera path exists,
-    the connection strategy below describes how to connect the edges of the two
+    For the case of $i = 0$, this would mean that the path does not exist. In
+other words, forward or backward path tracing, the latter of which isn't
+supported as camera lens are not defined in DTQ. For $i = 1$, it is either a
+point on the surface of a light, or the origin point on the 'delta eye'. Lastly,
+the BRDF is used to calculate the albedo of each primitive in the scene, and is
+then divided by the PDF sigma, which is the probability density of $v_i
+\rightarrow v_{i-1}$. $\alpha_1^L$ is shown below:
+
+\begin{lstlisting}
+  float3 N      = Sample_Cosine_Sphere(si, &pdf_pos),
+         origin = light.origin + light.radius*N,
+         dir    = Reorient_Angle(Sample_Cosine_Hemisphere(si, &pdf_dir), N);
+  float pdf_pos = 1.0f/(2.0f*TAU*SQR(light.radius)),
+        pdf_amt = 1.0f/EMITTER_AMT;
+  pv->irradiance = light.emission/(pdf_pos*pdf_amt*pdf_fwd);
+\end{lstlisting}
+
+  Note that the irradiance is being calculated in this case, that is, the amount
+  of flux arriving at point $pv$. Though in the case of $s=1$, this is more like
+  the radiance. The irradiance at $t > 1$ and $s > 1$ is shown:
+
+\begin{lstlisting}
+  float3 sample_dir, sample_f;
+  float sample_pdf;
+  _BSDF_Sample(wi, N, vtx, &sample_dir, &sample_pdf, &sample_f, si);
+  *irradiance = sample_f * (*irradiance) * fabs(dot(N, sample_dir))/sample_pdf;
+\end{lstlisting}
+
+  The connection strategy below describes how to connect the edges of the two
 paths

### 2017-11-29 dtoadqkernel.c

@@ -93,7 +93,7 @@ SampledPt SampledPt_From_Origin(float3 origin) {
 }
 
 typedef struct T_Vertex {
-  float3 origin, radiance, normal, mat_col;
+  float3 origin, irradiance, normal, mat_col;
   float pdf_fwd, pdf_bwd;
   Material* material;
   int M_ID;
@@ -257,6 +257,18 @@ float3 RColour ( float3 pt_colour, Material* m ) {
   return (pt_colour.x >= 0.0f) ? pt_colour : m->albedo;
 }
 
+bool Is_Delta ( Vertex* vtx ) {
+  return vtx->material == NULL ? 0 : ( vtx->material->specular     != 0.0f ||
+                                       vtx->material->transmittive != 0.0f);
+}
+
+bool Is_Black ( Vertex* vtx ) {
+  return vtx->irradiance.x == 0.0f &&
+         vtx->irradiance.y == 0.0f &&
+         vtx->irradiance.z == 0.0f;
+}
+
+
 float3 reflect ( float3 V, float3 N ) {
   return V - 2.0f*dot(V, N)*N;
 }
@@ -344,31 +356,33 @@ float Smith_G_GGX_Correlated ( float L, float R, float a ) {
 float3 BRDF_Diffuse_Sample ( float3 wi, float3 N, float* pdf, SceneInfo* si ) {
   return Reorient_Angle(Sample_Cosine_Hemisphere(si, pdf), N);
 }
-float BRDF_Diffuse_PDF  ( float3 wi, float3 N ) {
-  return PDF_Cosine_Hemisphere(wi, N);
+float BRDF_Diffuse_PDF  ( float3 wi, float3 wo, float3 N ) {
+  return PDF_Cosine_Hemisphere(wo, N);
 }
 float3 BRDF_Glossy_Sample ( float3 wi, float3 N, float* pdf, float glossy_lobe,
                             SceneInfo* si) {
   return Reorient_Angle(Sample_Uniform_Cone(glossy_lobe, pdf, si), N);
 }
-float BRDF_Glossy_PDF   ( float3 wi, float3 N, float cos_theta_max ) {
+float BRDF_Glossy_PDF   ( float3 wi, float3 wo, float3 N, float cos_theta_max ) {
   if ( cos_theta_max < 0.001f ) return 1.0f;
-  return PDF_Uniform_Cone(wi, N, cos_theta_max);
[…]

### 2017-11-29 testdocs/_region_.tex
whatever

@@
   dependent on $\alpha_{i-1}$, unless $i = 0$ in which case is the radiance of
-  either the emitter or the eye path. This is shown below.
+  either the emitter or the eye path. However, this can be simplified further,
+because of Helmhotz reciprocity, BRDFs return the same value if $wo$ and $wi$
+are switched. This is also the reason why forward and backward path tracing
+work.
   \begin{align}
@@
     \begin{cases}
-      \frac{L_e^0(y_0 \rightarrow y_1)}{p_A(y_1)} | \frac{W^0_e(z_0)}{p_A(z_0)},
-      &i = 1,\\
-      (\frac{f_s(y_{i+1} \rightarrow y_i \rightarrow y_{i-1})}{p_{\sigma}(y_i
-      \rightarrow y_{i-1})} | \frac{f_s(z_{i-1} \rightarrow z_{i-2} \rightarrow
-      z_{i-3}}{p_\sigma(z_{i-2} \rightarrow z_{i-1})}) \alpha^{L|E}_{i-1},
-      &i > 1
+      1.0 &i = 0,\\
+      \frac{
+      L_e^0(V_0 \rightarrow V_1)}{p_A(V_1)} | 1.0 &i = 1,\\
+      \frac{f_s(V_0 \rightarrow V_1 \rightarrow V_2)}{p_{\sigma}(V_0
+      \rightarrow V_1)} &i > 1
     \end{cases}
@@
 other words, forward or backward path tracing, the latter of which isn't
-supported as camera lens are not defined in DTQ. Veach (and many other sources),
-leave this in, but it's unnecessarily confusing. For $i = 1$, it is either a
-point on the surface of a light, or the origin point on the 'delta eye'. The
-connection strategy below describes how to connect the edges of the two paths
+supported as camera lens are not defined in DTQ. For $i = 1$, it is either a
+point on the surface of a light, or the origin point on the 'delta eye'. Lastly,
+the BRDF is used to calculate the albedo of each primitive in the scene, and is
+then divided by the PDF sigma, which is the probability density of $v_i
+\rightarrow v_{i-1}$. $\alpha_1^L$ is shown below:
+
+\begin{lstlisting}
+  float3 N      = Sample_Cosine_Sphere(si, &pdf_pos),
+         origin = light.origin + light.radius*N,
+         dir    = Reorient_Angle(Sample_Cosine_Hemisphere(si, &pdf_dir), N);
+  float pdf_pos = 1.0f/(2.0f*TAU*SQR(light.radius)),
+        pdf_amt = 1.0f/EMITTER_AMT;
+  pv->irradiance = light.emission/(pdf_pos*pdf_amt*pdf_fwd);
+\end{lstlisting}
+
+  Note that the irradiance is being calculated in this case, that is, the amount
+  of flux arriving at point $pv$. Though in the case of $s=1$, this is more like
+  the radiance. The irradiance at $t > 1$ and $s > 1$ is shown:
+
+\begin{lstlisting}
+  float3 sample_dir, sample_f;
+  float sample_pdf;
+  _BSDF_Sample(wi, N, vtx, &sample_dir, &sample_pdf, &sample_f, si);
+  *irradiance = sample_f * (*irradiance) * fabs(dot(N, sample_dir))/sample_pdf;
+\end{lstlisting}
+
+  The connection strategy below describes how to connect the edges of the two
+paths. There is no $t = 0$ case as no physical camera lens exists on the scene.
   \begin{align}
@@
     \begin{cases}
-      L_e(z_{-2} \rightarrow z_{-1}), &s = 0, t > 1,\\
-      f_s(y_{-2} \rightarrow y_{-1} \rightarrow z_{-1}) G(y_{-1} \leftrightarrow
-      z_{-1}) f_s(y_{-1} \rightarrow z_{-1} \rightarrow z_{-2}), &s > 0, t > 1
+      L_e(L0 \rightarrow L1), &s = 0, t > 1,\\
+      f_s(L0 \rightarrow L1 \rightarrow L2) G(E1 \leftrightarrow
+      L1) f_s(L1 \rightarrow E1 \rightarrow E0), &s > 0, t > 1
     \end{cases}
   \end{align}
+  In code:
+  \begin{lstlisting}
+Spectrum contribution = E1.irradiance * L1->irradiance;
+
+// As there is no T=1 case, L1 -> E1 -> E0 [Le in s=1 case] must happen
+contribution *= Subpath_Connection(L1, &E1, &E0);
+
+// For s == 1, L1 = L0, a point on the surface of emitter, there is no
+// L0 -> L1 -> E1 connection
+if ( light_depth > 0 ) {
+  // Connect L0 -> L1 -> E1
+  contribution *= Subpath_Connection( L0,  L1, &E1);
+}
+
+// Geometric connection term [includes visibility check]
+contribution *= Geometric_Term(&E1, L1, si, Tx);
+
+  \end{lstlisting}
+
     Is the $s = 0, t > 1$ strategy worthwhile to compute? Those familiar with
@@
 pointless anyways, as a path generated by light-tracing would have to be able to
-contribute to any pixel, which is not feasible on a GPU. The case of $t = 1$,
-where the light path is connected to a position on the lens, is important for
-effects such as depth of field and aperture. The only problem is due to time
-constraints, this has not been implemented yet.
-
-    One immediately obvious optimization that could be made with this model is
-in regards to the visibility check in the geometric term, if you were to expand
-$I_j$, and take the V term from outside the G function
-    [$G(v \rightarrow
-    \dot{v}) V(v \rightarrow \dot{v})$],
-    it would be made obvious that all the visibility checks can be cancelled out
-for the entire equation, except for the connection term $c$. This is intuitive
-as the paths $\bar{y}_{0} \ldots \bar{y}_s$ and $\bar{z}_0 \ldots \bar{z}_t$ had
-just been generated via raymarching. Specifically, the definition of $V$ is
-  \begin{align}
-    V(v \rightarrow \dot{v}) = ||v - \dot{v}|| \leq
-    m(v,\:\overrightarrow{v\:\dot{v}})
-  \end{align}
-    where $m$ is a march through the SDF map. A visibility check is only
-necessary for the connection edge
-    \begin{align*}
-      ||v - \dot{v}|| = m(v, \overrightarrow{v, \dot{v}})\\
-      \:\:\:\text{for all paths but } y_{-1} \rightarrow z_{-1} 
-    \end{align*}
+contribute to any pixel, which is not feasible on a GPU.
 \\
@@
     where $R$ is the radius of the sphere, note $y_0$ is a point on the surface
-of the sphere, and not its origin. \\$p_A^E(x_0 \rightarrow x_1)$ doesn't exist,
-as there is no physical lens on the scene. [TODO - this might cahnge later]. One
-of the benefits of this, is that there are no special cases for pathes where $t
-= 0$, which is impossible to handle on the GPU as the path could contribute to
-any arbitrary pixel [directly hitting the lens].
+    of the sphere, and not its origin. For the eye:
   \begin{align}
@@
 \end{document}
-\message{ !name(test.tex) !offset(-684) }
+\message{ !name(test.tex) !offset(-707) }

### 2017-11-29 testdocs/test.tex
whatever

@@
       \frac{
-      L_e^0(y_0 \rightarrow y_1)}{p_A(y_1)} | 1.0 &i = 1,\\
-      \frac{f_s(v_{i-1} \rightarrow v_i \rightarrow v_{i+1})}{p_{\sigma}(v_i
-      \rightarrow v_{i-1})} &i > 1
+      L_e^0(V_0 \rightarrow V_1)}{p_A(V_1)} | 1.0 &i = 1,\\
+      \frac{f_s(V_0 \rightarrow V_1 \rightarrow V_2)}{p_{\sigma}(V_0
+      \rightarrow V_1)} &i > 1
     \end{cases}
@@
   The connection strategy below describes how to connect the edges of the two
-paths
+paths. There is no $t = 0$ case as no physical camera lens exists on the scene.
   \begin{align}
@@
     \begin{cases}
-      L_e(z_{-2} \rightarrow z_{-1}), &s = 0, t > 1,\\
-      f_s(y_{-2} \rightarrow y_{-1} \rightarrow z_{-1}) G(y_{-1} \leftrightarrow
-      z_{-1}) f_s(y_{-1} \rightarrow z_{-1} \rightarrow z_{-2}), &s > 0, t > 1
+      L_e(L_0 \rightarrow L_1), &s = 0, t > 1,\\
+      f_s(L_0 \rightarrow L_1 \rightarrow L2) G(E_1 \leftrightarrow
+      L_1) f_s(L_1 \rightarrow E_1 \rightarrow E_0), &s > 0, t > 1
     \end{cases}
   \end{align}
+  In code:
+  \begin{lstlisting}
+Spectrum contribution = E1.irradiance * L1->irradiance;
+
+// As there is no T=1 case, L1 -> E1 -> E0 [Le in s=1 case] must happen
+contribution *= Subpath_Connection(L1, &E1, &E0);
+
+// For s == 1, L1 = L0, a point on the surface of emitter, there is no
+// L0 -> L1 -> E1 connection
+if ( light_depth > 0 ) {
+  // Connect L0 -> L1 -> E1
+  contribution *= Subpath_Connection( L0,  L1, &E1);
+}
+
+// Geometric connection term [includes visibility check]
+contribution *= Geometric_Term(&E1, L1, si, Tx);
+
+  \end{lstlisting}
+
     Is the $s = 0, t > 1$ strategy worthwhile to compute? Those familiar with
@@
 pointless anyways, as a path generated by light-tracing would have to be able to
-contribute to any pixel, which is not feasible on a GPU. The case of $t = 1$,
-where the light path is connected to a position on the lens, is important for
-effects such as depth of field and aperture. The only problem is due to time
-constraints, this has not been implemented yet.
+contribute to any pixel, which is not feasible on a GPU.
+\\
 
-    One immediately obvious optimization that could be made with this model is
-in regards to the visibility check in the geometric term, if you were to expand
-$I_j$, and take the V term from outside the G function
-    [$G(v \rightarrow
-    \dot{v}) V(v \rightarrow \dot{v})$],
-    it would be made obvious that all the visibility checks can be cancelled out
-for the entire equation, except for the connection term $c$. This is intuitive
-as the paths $\bar{y}_{0} \ldots \bar{y}_s$ and $\bar{z}_0 \ldots \bar{z}_t$ had
-just been generated via raymarching. Specifically, the definition of $V$ is
-  \begin{align}
-    V(v \rightarrow \dot{v}) = ||v - \dot{v}|| \leq
-    m(v,\:\overrightarrow{v\:\dot{v}})
-  \end{align}
-    where $m$ is a march through the SDF map. A visibility check is only
-necessary for the connection edge
-    \begin{align*}
-      ||v - \dot{v}|| = m(v, \overrightarrow{v, \dot{v}})\\
-      \:\:\:\text{for all paths but } y_{-1} \rightarrow z_{-1} 
-    \end{align*}
-\\
-    Moving on, for $\alpha^{L|E}_i$, these are equivalent for both $L$ and $E$,
-the problem is that the $\bar{y}$ path is generated in perspective of the light
-source, so there has to be a bit of caution in terms of perspective in when
-generating the PDFs. $p_A$ is the PDF w.r.t. area, for the light path; we only
-need to take the PDF of the solid angle of a sphere
-  \begin{align}
-    p_A^L(y_0 \rightarrow y_1) = p_{\sigma}^L(y_0 \rightarrow y_1) G(v
-    \leftrightarrow \dot{v})\\
-    p_{\sigma}^L(y_0 \rightarrow y_1) = \frac{1.0f}{\tau * (1.0f -
-    \sqrt{\frac{R^2}{|y_0 - y_1|^2}})}
-  \end{align}
-    where $R$ is the radius of the sphere, note $y_0$ is a point on the surface
-of the sphere, and not its origin. \\$p_A^E(x_0 \rightarrow x_1)$ doesn't exist,
-as there is no physical lens on the scene. [TODO - this might cahnge later]. One
-of the benefits of this, is that there are no special cases for pathes where $t
-= 0$, which is impossible to handle on the GPU as the path could contribute to
-any arbitrary pixel [directly hitting the lens].
-  \begin{align}
-    P_A^E = 1.0f
-  \end{align}
-    The $L^0_e$ is contribution from the light source. As of right now, DTOADQ
-only supports area lights, so $L^0_e$ is set a constant for the emitter's
-spectrum value. The $W_e^0$ is contribution from the eye source, that is, the
-lens. Since the lens does not exist in the scene, as of now [TODO - this might
-change later], there is no $W_e^0$. While this should be implemented in the
-future, a hidden benefit of the current model is that there are no special cases
-for paths where $t = 0$, which is impossible to handle on the GPU as any such
-path could contribute to any arbitrary pixel [the light path is, after all,
-directly hitting the lens]. So this has now become:
-  \begin{align}
-    \alpha_i^E = \Sigma_{i\leq 0} 
-  \end{align}
-  The amount that each path contributes to F differ greatly, so naively weighing
+The amount that each path contributes differ greatly, so naively weighing
 each contribution uniformly by something like 1.0f, or even, by inverse path
 length, will producy very noisy images, and would not be worth the additional
-computation time that BDPT requires, as shown below. To calculate the PDF of In
-order to get good samples from the contributions, they must be combined in an
-optimal manner using Multiple Importance Sampling: each contribution is weighted
-by the PDF of the entire path itself:
-  \begin{align*}
-    F &\equiv \Sigma_s\Sigma_tW_{s, t}C_{s, t}\\
-    C_{s, t} &= \alpha^L_s c_{s, t} \alpha^E_t
-  \end{align*}
-    where $W_{s, t}$ is the MIS weight of path $\bar{y} \leftarrow \bar{z}$, and
-$C_{s, t}$ is the unweighted contribution of the same path. For clarity,
-$C^U_{s, t}$ as described below is the unweighted contribution of the
-    algorithm:
-  \begin{align}
-    C^u_{s, t} = (\alpha^L_s c_{s, t} \alpha^E_t) \equiv
-    \frac{f_j(\bar{v})}{p(\bar{v})}
-  \end{align}
-  thus
-  \begin{align}
-    F = \Sigma_s\Sigma_tW_{s, t}C^u_{s, t}
-  \end{align}
-    The formula for Multiple Importance Sampling on BDPT is as so:
+computation time that BDPT requires. Thus, as mentioned earlier, probability
+distributions of the entire path are taken into account in order to generate
+multiple importance samples.
+    The formula for multiple importance sampling on BDPT is as so:
   \begin{align}
@@
 generated, while the denominator is the path density of other connection
-strategies that could could have, in theory,$^{[1]}$ created the path. The
+strategies that could could have, in theory, created the path. The
 straight forward implementation of this has a lot of problems gone unresolved in
@@
 contribution is inversed from this, (also iterating towards s, hence why
-$\bar{x}_{i-1}$ is used). Another way of saying it, is $\bar{x} = \bar{y} ||
-\bar{z}_{rev}$. In DTOADQ's case, the weights are calculated in the same
-direction relative to its originator; in this case the perspective of the eye
-path.
+$\bar{x}_{i-1}$ is used). . In DTOADQ's case, the weights are calculated in the
+same direction relative to its originator; in this case the perspective of the
+eye path. Not only this, but unlike any implementation I've found of BDPT, I
+don't recalculate each path's probability and MIS every vertex. That is, usually
+on calculating MIS for each connection, the MIS and probability values are
+recalculated for each vertex in the path. As this is not possible given that
+only the light-path exists in memory, an array for MIS/probability is stored
+corresponding to each light vertex, and a single MIS/probability is stored
+corresponding to $E_1$. Thus, the weight on each path looks like:
   \begin{align}

### 2017-11-29 dtoadqkernel.c

@@ -43,7 +43,7 @@ bool Is_Debug ( ) {
 typedef struct T_Camera {
   float3 position, lookat, up;
   int2 dim;
-  float fov;
+  float fov, focal;
   int flags;
 } Camera;
 
@@ -865,12 +865,14 @@ Spectrum BDPT_Integrate ( float3 pixel, float3 dir, SCENE_T(si, Tx)) {
 
       Spectrum contribution = E1.irradiance * L1->irradiance;
 
-      if ( light_depth > 0 ) { // s >= 1, t >= 1 strategy
-        contribution *= Subpath_Connection(&E0, &E1, L1);
+      // As there is no T=1 case, L1 -> E1 -> E0 [Le in s=1 case] must happen
+      contribution *= Subpath_Connection(L1, &E1, &E0);
+
+      // For s == 1, L1 = L0, a point on the surface of emitter, there is no
+      // L0 -> L1 -> E1 connection
+      if ( light_depth > 0 ) {
+        // Connect L0 -> L1 -> E1
         contribution *= Subpath_Connection( L0,  L1, &E1);
-      } else { // s == 1 strategy. L1 = L0, a point on the surface of area light
-        // Connect E0 -> E1 -> L1
-        contribution *= Subpath_Connection(&E0, &E1, L1);
       }
 
       // Geometric connection term [includes visibility check]

### 2017-11-29 raycastkernel.c

@@ -27,7 +27,7 @@ __constant float MARCH_ACC = //%MARCH_ACC.0f/1000.0f;
 typedef struct T_Camera {
   float3 position, lookat, up;
   int2 dim;
-  float fov;
+  float fov, focal;
   int flags;
 } Camera;
 
@@ -76,6 +76,29 @@ typedef struct T_SampledPt {
   float dist;
   int material_index;
 } SampledPt;
+
+// -----------------------------------------------------------------------------
+// --------------- GENERAL FUNCTIONS      --------------------------------------
+float Rand ( SceneInfo* si ) {
+  enum { A=4294883355U };
+  uint2 r = (*si).rng_state;
+  uint res = r.x^r.y;
+  uint hi = mul_hi(r.x, A);
+  r.x = r.x*A + r.y;
+  r.y = hi + (r.x<r.y);
+  (*si).rng_state = r;
+  return res/(float)(UINT_MAX);
+}
+
+float Sample_Uniform ( SceneInfo* si ) {
+  return Rand(si);
+}
+float2 Sample_Uniform2 ( SceneInfo* si ) {
+  return (float2)(Sample_Uniform(si), Sample_Uniform(si));
+}
+float3 Sample_Uniform3 ( SceneInfo* si ) {
+  return (float3)(Sample_Uniform(si), Sample_Uniform2(si));
+}
 // -----------------------------------------------------------------------------
 // --------------- MAP GEOMETRY FUNCTIONS --------------------------------------
 
@@ -149,10 +172,12 @@ float3 RColour ( float3 pt_colour, Material* m ) {
 
 // -----------------------------------------------------------------------------
 // --------------- CAMERA ------------------------------------------------------
-Ray Camera_Ray(Camera* camera) {
+Ray Camera_Ray(Camera* camera, SceneInfo* si) {
   float2 coord = (float2)((float)get_global_id(0), (float)get_global_id(1));
   float2 resolution = (float2)((float)camera->dim.x, (float)camera->dim.y);
   resolution.y *= 16.0f/9.0f;
+
+  float fov_r = (180.0f - camera->fov)*PI/180.0f;
 
   float2 mouse_pos = camera->lookat.xy;
 
@@ -168,9 +193,17 @@ Ray Camera_Ray(Camera* camera) {
 
   float3 cam_up = normalize(cross(cam_right, cam_front));
[…]

### 2017-11-29 testdocs/_region_.tex
doicumentation

@@
 hand allows a cheap normal computation of primitives such as a spheres.
+
+  Overall, it seems that while SDFs are fun to play around with, for the case of
+  writing a generalized light transport algorithm to apply to SDFs causes a lot
+  of problems while being generally limiting. The biggest problem is dealing
+  with transparent materials, and dealing with visibility checks. For
+  transparent materials, it's required that the ray march 'through' the SDF
+  primitive, and find the exit. The solution is to apply a 'shell' to the SDF,
+  or in other words, force the SDF to be unsigned:
+
+\begin{lstlisting}
+float Shell ( float dist, float r ) {
+  return fabs(dist) - r*0.5f;
+}
+\end{lstlisting}
+
   \section{OpenCL}
 	OpenCL, and GPU programming in general, is where I am least knowledgeable or
-interested. Well, OpenCL is easy to work with, and I wrote my own OpenCL module
+interested. Well, OpenCL is easy enough to work with compared to OGL or CUDA,
+however I'm not good at dealing with the intrinsics of the GPU: optimizing code,
+knowing how memory and thread/workers should be handled, etc. For example,
+probably the biggest problem with the DTQ kernel right now is it is a single
+file megakernel. There just was never any interest to split it into multiple
+files nor multiple kernels. However, working directly with the OpenCL api is
+easy enough, and I did write my own OpenCL module
 to handle OCL in a rather convenient method, for example, the DTOADQ kernel call
@@
     \begin{cases}
-      L_e(L0 \rightarrow L1), &s = 0, t > 1,\\
-      f_s(L0 \rightarrow L1 \rightarrow L2) G(E1 \leftrightarrow
-      L1) f_s(L1 \rightarrow E1 \rightarrow E0), &s > 0, t > 1
+      L_e(L_0 \rightarrow L_1), &s = 0, t > 1,\\
+      f_s(L_0 \rightarrow L_1 \rightarrow L2) G(E_1 \leftrightarrow
+      L_1) f_s(L_1 \rightarrow E_1 \rightarrow E_0), &s > 0, t > 1
     \end{cases}
@@
 \\
-    Moving on, for $\alpha^{L|E}_i$, these are equivalent for both $L$ and $E$,
-the problem is that the $\bar{y}$ path is generated in perspective of the light
-source, so there has to be a bit of caution in terms of perspective in when
-generating the PDFs. $p_A$ is the PDF w.r.t. area, for the light path; we only
-need to take the PDF of the solid angle of a sphere
-  \begin{align}
-    p_A^L(y_0 \rightarrow y_1) = p_{\sigma}^L(y_0 \rightarrow y_1) G(v
-    \leftrightarrow \dot{v})\\
-    p_{\sigma}^L(y_0 \rightarrow y_1) = \frac{1.0f}{\tau * (1.0f -
-    \sqrt{\frac{R^2}{|y_0 - y_1|^2}})}
-  \end{align}
-    where $R$ is the radius of the sphere, note $y_0$ is a point on the surface
-    of the sphere, and not its origin. For the eye:
-  \begin{align}
-    P_A^E = 1.0f
-  \end{align}
-    The $L^0_e$ is contribution from the light source. As of right now, DTOADQ
-only supports area lights, so $L^0_e$ is set a constant for the emitter's
-spectrum value. The $W_e^0$ is contribution from the eye source, that is, the
-lens. Since the lens does not exist in the scene, as of now [TODO - this might
-change later], there is no $W_e^0$. While this should be implemented in the
-future, a hidden benefit of the current model is that there are no special cases
-for paths where $t = 0$, which is impossible to handle on the GPU as any such
-path could contribute to any arbitrary pixel [the light path is, after all,
-directly hitting the lens]. So this has now become:
-  \begin{align}
-    \alpha_i^E = \Sigma_{i\leq 0} 
-  \end{align}
-  The amount that each path contributes to F differ greatly, so naively weighing
+
+The amount that each path contributes differ greatly, so naively weighing
 each contribution uniformly by something like 1.0f, or even, by inverse path
 length, will producy very noisy images, and would not be worth the additional
-computation time that BDPT requires, as shown below. To calculate the PDF of In
-order to get good samples from the contributions, they must be combined in an
-optimal manner using Multiple Importance Sampling: each contribution is weighted
-by the PDF of the entire path itself:
-  \begin{align*}
-    F &\equiv \Sigma_s\Sigma_tW_{s, t}C_{s, t}\\
-    C_{s, t} &= \alpha^L_s c_{s, t} \alpha^E_t
-  \end{align*}
-    where $W_{s, t}$ is the MIS weight of path $\bar{y} \leftarrow \bar{z}$, and
-$C_{s, t}$ is the unweighted contribution of the same path. For clarity,
-$C^U_{s, t}$ as described below is the unweighted contribution of the
-    algorithm:
-  \begin{align}
-    C^u_{s, t} = (\alpha^L_s c_{s, t} \alpha^E_t) \equiv
-    \frac{f_j(\bar{v})}{p(\bar{v})}
-  \end{align}
-  thus
-  \begin{align}
-    F = \Sigma_s\Sigma_tW_{s, t}C^u_{s, t}
-  \end{align}
-    The formula for Multiple Importance Sampling on BDPT is as so:
+computation time that BDPT requires. Thus, as mentioned earlier, probability
+distributions of the entire path are taken into account in order to generate
+multiple importance samples.
+    The formula for multiple importance sampling on BDPT is as so:
   \begin{align}
@@
   \end{align}
+
+  The camera handling function is based off glLookAt. The difference is that it
+  supports field of view, depth of field and antialiasing. The model isn't
+perfect and has a few bugs, mostly because the camera used in DTQ
+is a pinhole camera, or as I call it delta camera. In general, very little time
+was spent on the camera mechanics. In the future, there will be a physical
+camera lens in the SDF, so as to actually simulate depth of field, aperture,
+etc. Shown below is an exaggerated depth-of-field in the raytracer.
+\\
+
+  \includegraphics[scale=0.5]{rt_dof}
 
@@
 \end{document}
-\message{ !name(test.tex) !offset(-707) }
+\message{ !name(test.tex) !offset(-703) }

### 2017-11-29 testdocs/test.tex
doicumentation

@@
 
+  The camera handling function is based off glLookAt. The difference is that it
+  supports field of view, depth of field and antialiasing. The model isn't
+perfect and has a few bugs, mostly because the camera used in DTQ
+is a pinhole camera, or as I call it delta camera. In general, very little time
+was spent on the camera mechanics. In the future, there will be a physical
+camera lens in the SDF, so as to actually simulate depth of field, aperture,
+etc. Shown below is an exaggerated depth-of-field in the raytracer.
+\\
+
+  \includegraphics[scale=0.5]{rt_dof}
+
   \section{DTOADQ BRDF}

### 2017-11-29 dtoadqkernel.c

@@ -1027,7 +1027,7 @@ __kernel void DTOADQ_Kernel (
     _EvalPreviousOutputHorcrux _hor =
           Eval_Previous_Output(img, output_img, out, sinfo, pix_pt);
     old_colour = _hor.old_colour;
-    if ( old_colour.w >= spp ) return;
+    if ( old_colour.w >= 64 ) return;
     if ( _hor.raycast_nav && DO_NAVIGATION ) {
       // Laughably bad raycast to make it easy to navigate a scene
       SampledPt pt = March(-1, ray, &scene_info, textures);

### 2017-11-29 raycastkernel.c

@@ -1,6 +1,7 @@
 #define MAX_DEPTH 16
 #define MARCH_DIST //%MARCH_DIST.0f
 #define MARCH_REPS //%MARCH_REPS
+#define MAT_LEN //%MAT_LENGTH
 
 #define TEXTURE_T __read_only image2d_array_t
 #define SCENE_T(S, T) SceneInfo* S , TEXTURE_T T
@@ -27,7 +28,7 @@ __constant float MARCH_ACC = //%MARCH_ACC.0f/1000.0f;
 typedef struct T_Camera {
   float3 position, lookat, up;
   int2 dim;
-  float fov, focal;
+  float fov, focal, radius;
   int flags;
 } Camera;
 
@@ -46,20 +47,27 @@ typedef struct T_Emitter {
   float radius;
 } Emitter;
 
+typedef struct T_SharedInfo {
+  unsigned char clear_img;
+  unsigned long finished_samples;
+  unsigned char spp;
+  uint2 rng_state;
+} SharedInfo;
+
 typedef struct T_SceneInfo {
   float time;
   // __read_only image2d_array_t textures; IMAGES CANT BE USED AS FIELD TYPES:-(
-  __global Material* materials;
+  Material* materials;
   float3 debug_values;
-  __global int* rng;
+  uint2 rng_state;
 } SceneInfo;
-SceneInfo New_SceneInfo(float time, __global Material* materials,
-                        float3 debug_values, __global int* rng){
+SceneInfo New_SceneInfo(float time, Material* materials,
+                        float3 debug_values, uint2 rng){
   SceneInfo si;
   si.time         = time;
   si.materials    = materials;
   si.debug_values = debug_values;
-  si.rng          = rng;
+  si.rng_state    = rng;
   return si;
 }
 
@@ -79,6 +87,7 @@ typedef struct T_SampledPt {
 
 // -----------------------------------------------------------------------------
 // --------------- GENERAL FUNCTIONS      --------------------------------------
+
 float Rand ( SceneInfo* si ) {
   enum { A=4294883355U };
   uint2 r = (*si).rng_state;
@@ -95,6 +104,10 @@ float Sample_Uniform ( SceneInfo* si ) {
 }
 float2 Sample_Uniform2 ( SceneInfo* si ) {
   return (float2)(Sample_Uniform(si), Sample_Uniform(si));
+}
+// -1 .. 1
+float2 Sample_Uniform2_2 ( SceneInfo* si ) {
[…]

### 2017-11-29 raytracekernel.c

@@ -34,9 +34,10 @@ bool Is_Debug ( ) {
 typedef struct T_Camera {
   float3 position, lookat, up;
   int2 dim;
-  float fov;
+  float fov, focal, radius;
   int flags;
 } Camera;
+
 
 typedef struct T_Material {
   // colour [set to (-1.0, -1.0, -1.0) to have map override it]
@@ -53,18 +54,27 @@ typedef struct T_Emitter {
   float radius;
 } Emitter;
 
+typedef struct T_SharedInfo {
+  unsigned char clear_img;
+  unsigned long finished_samples;
+  unsigned char spp;
+  uint2 rng_state;
+} SharedInfo;
+
 typedef struct T_SceneInfo {
   float time;
   // __read_only image2d_array_t textures; IMAGES CANT BE USED AS FIELD TYPES:-(
   Material* materials;
   float3 debug_values;
+  uint2 rng_state;
 } SceneInfo;
 SceneInfo New_SceneInfo(float time, Material* materials,
-                        float3 debug_values){
+                        float3 debug_values, uint2 rng){
   SceneInfo si;
   si.time         = time;
   si.materials    = materials;
   si.debug_values = debug_values;
+  si.rng_state    = rng;
   return si;
 }
 
@@ -84,6 +94,38 @@ typedef struct T_SampledPt {
   float dist;
   int material_index;
 } SampledPt;
+
+float Rand ( SceneInfo* si ) {
+  enum { A=4294883355U };
+  uint2 r = (*si).rng_state;
+  uint res = r.x^r.y;
+  uint hi = mul_hi(r.x, A);
+  r.x = r.x*A + r.y;
+  r.y = hi + (r.x<r.y);
+  (*si).rng_state = r;
+  return res/(float)(UINT_MAX);
+}
+
+float3 Float3_Max(float3 t, float g) {
+  return (float3)(fmax(t.x, g), fmax(t.y, g), fmax(t.z, g));
+}
+float3 Float3_Min(float3 t, float g) {
+  return (float3)(fmin(t.x, g), fmin(t.y, g), fmin(t.z, g));
+}
+
+float Sample_Uniform ( SceneInfo* si ) {
+  return Rand(si);
+}
+float2 Sample_Uniform2 ( SceneInfo* si ) {
+  return (float2)(Sample_Uniform(si), Sample_Uniform(si));
+}
+// -1 .. 1
+float2 Sample_Uniform2_2 ( SceneInfo* si ) {
[…]

### 2017-12-04 testdocs/test.tex
'finished' BDPT got transmittive to work

@@
   \maketitle
-  \section{Introduction}
-
   \section{Signed Distance Fields}
@@
 
-  This is a larger problem in Light Transport than it is in raytracing because
+  This is a larger problem in light transport than it is in raytracing because
 small discontinuities can impact the lighting of an entire scene. And to add to
@@
 recalculated for each vertex in the path. As this is not possible given that
-only the light-path exists in memory, an array for MIS/probability is stored
-corresponding to each light vertex, and a single MIS/probability is stored
+only the light-path exists in memory, an array for sum/probability is stored
+corresponding to each light vertex, and a single sum/probability is stored
 corresponding to $E_1$. Thus, the weight on each path looks like:
@@
   \end{align}
-    And then the expanded form of $\mathcal{W}_i(\bar{x})$ looks like
-  \begin{align}
-    \mathcal{W}_i(\bar{x}) = \Pi_{n=2}^i(\frac{\overleftarrow
-    {p_{\sigma}}(x_n)\overleftarrow {G}(x_n)}
-    {\overrightarrow{p_{\sigma}}(x_n)\overrightarrow{G}(x_n)})
-    * \frac{\overleftarrow{p_A}(x_1)}{\overrightarrow{p_A}(x_1)}
-  \end{align}
     And now the MIS looks like:
@@
     \frac{1.0f}{\mathcal{W}_s(\bar{y}) + \mathcal{W}_t(\bar{z}) + 1.0f}
-  \end{align}
-    Thus, the final rendering equation for DTOADQ looks like such:
-  \begin{align}
-    P = \frac{1.0f}{N} \Sigma_{0}^{N} \Sigma_{s \le 0} \Sigma_{t \le 0}
-    \alpha^L_s c_{s,t} \alpha^E_t MIS_{s, t}(\bar{y}, \bar{z})
-  \end{align}
-
-  \begin{align}
-    P_i(\bar{x}_{s, t}) = P^L_s p^E_t
-  \end{align}
-
-  \begin{align}
-    p^{L|E}_i(\bar{x}) =
-    \begin{cases}
-      1.0f, & i = 0,\\
-      P_A(x_i), & i = 1,\\
-      P_{\sigma}(x_{i-1} \rightarrow x_i) G(x_{i-1} \leftrightarrow x_i)
-      P^{L|E}_{i-1}, & i > 1
-    \end{cases}
   \end{align}

### 2017-12-04 dtoadqkernel.c

@@ -1,4 +1,4 @@
-#define MAX_DEPTH 5
+#define MAX_DEPTH 4
 #define MARCH_DIST //%MARCH_DIST.0f
 #define MARCH_REPS //%MARCH_REPS
 #define MAT_LEN //%MAT_LENGTH
@@ -7,7 +7,8 @@
 #define SCENE_T(S, T) SceneInfo* S, TEXTURE_T T
 // EMIT_MAT = -120 - light_index
 #define EMIT_MAT -120
-#define REMIT(I) REmission(EMIT_MAT - I, si->debug_values, si->time)
+#define REMITM(I) (EMIT_MAT - (I))
+#define REMIT(I) REmission(REMITM(I), si->debug_values, si->time)
 
 #define DFLT2(V) ((V).x), ((V).y)
 #define DFLT3(V) ((V).x), ((V).y), ((V).z)
@@ -43,7 +44,7 @@ bool Is_Debug ( ) {
 typedef struct T_Camera {
   float3 position, lookat, up;
   int2 dim;
-  float fov, focal;
+  float fov, focal, radius;
   int flags;
 } Camera;
 
@@ -60,6 +61,7 @@ typedef struct T_Material {
 typedef struct T_Emitter {
   float3 origin, emission;
   float radius;
+  int index;
 } Emitter;
 
 typedef struct T_SharedInfo {
@@ -100,7 +102,11 @@ typedef struct T_Vertex {
 } Vertex;
 
 float Calc_Prob(Vertex* t) {
-  return t->pdf_bwd/t->pdf_fwd;
+  float pdf_bwd = t->pdf_bwd,
+        pdf_fwd = t->pdf_fwd;
+  if ( pdf_bwd == 0.0f ) pdf_bwd = 1.0f;
+  if ( pdf_fwd == 0.0f ) pdf_fwd = 1.0f;
+  return pdf_bwd/pdf_fwd;
 }
 
 typedef struct T_Subpath {
@@ -155,6 +161,10 @@ float Sample_Uniform ( SceneInfo* si ) {
 float2 Sample_Uniform2 ( SceneInfo* si ) {
   return (float2)(Sample_Uniform(si), Sample_Uniform(si));
 }
+// -1 .. 1
+float2 Sample_Uniform2_2 ( SceneInfo* si ) {
+  return ((float2)(Sample_Uniform(si), Sample_Uniform(si))-(float2)(0.5f))*2.0f;
+}
[…]

### 2017-12-04 raytracekernel.c

@@ -414,8 +414,6 @@ Ray Camera_Ray(Camera* camera, SceneInfo* si) {
   float3 cam_right = normalize ( cross(cam_front, (float3)(0.0f, 1.0f, 0.0f)));
 
   float3 cam_up = normalize(cross(cam_right, cam_front));
-  Ray ray;
-
   /* // ------ DOF & antialiasing ----- */
   // Adapted from https://www.shadertoy.com/view/lsX3DH
   float3 dof_puv = (float3)(puv, fov_r);
@@ -425,9 +423,7 @@ Ray Camera_Ray(Camera* camera, SceneInfo* si) {
   cam_pos += dof_origin.x*cam_right + dof_origin.y*cam_up;
   ray_dir += dof_dir.x*cam_right + dof_dir.y*cam_up;
   ray_dir = normalize(ray_dir);
-  ray.origin = cam_pos;
-  ray.dir = ray_dir;
-  return ray;
+  return (Ray){cam_pos, ray_dir};
 }
 
 typedef struct _T_EvalPreviousOutputHorcrux {

### 2017-12-05 dtoadqkernel.c
glossy

@@ -332,7 +332,7 @@ float PDF_Cosine_Sphere ( float3 wi ) {
 }
 
 float _PDF_Uniform_Cone ( float cos_theta_max ) {
-  return 4.0f*PI*sqr(sin(0.5f*cos_theta_max));
+  return 1.0f/(4.0f*PI*sqr(sin(0.5f*cos_theta_max)));
 }
 float3 Sample_Uniform_Cone ( float cos_theta_max, float* pdf, SceneInfo* si ) {
   float2 u = Sample_Uniform2(si);
@@ -379,7 +379,8 @@ float BRDF_Diffuse_PDF  ( float3 wi, float3 wo, float3 N ) {
 }
 float3 BRDF_Glossy_Sample ( float3 wi, float3 N, float* pdf, float glossy_lobe,
                             SceneInfo* si) {
-  return Reorient_Angle(Sample_Uniform_Cone(glossy_lobe, pdf, si), N);
+  float3 NN = reflect(wi, N);
+  return Reorient_Angle(Sample_Uniform_Cone(glossy_lobe, pdf, si), NN);
 }
 float BRDF_Glossy_PDF   ( float3 wi, float3 wo, float3 N, float cos_theta_max ) {
   if ( cos_theta_max < 0.001f ) return 1.0f;
@@ -415,14 +416,14 @@ Spectrum _BRDF_F ( float3 wi, float3 N, float3 wo, Material* m, float3 col ) {
                Fresnel_L = Schlick_Fresnel(cos_NL),
                Fresnel_V = Schlick_Fresnel(cos_NV);
   // Diffusive component
-  float3 diffusive_albedo = col * IPI;
+  float3 diffusive_albedo = col*IPI;
 
   float3 microfacet = (float3)(1.0f);
 
 
   // probably transmittive
   if ( cos_NL <= 0.0f || cos_NV <= 0.0f ) {
-    diffusive_albedo *= PI*1.0f*m->transmittive;
+    diffusive_albedo *= PI*m->transmittive;
     return diffusive_albedo;
   }
 
@@ -610,7 +611,7 @@ SampledPt March_Visibility ( int avoid, Ray ray, SCENE_T(si, Tx)) {
 float Visibility_Ray(float3 orig, float3 other, SCENE_T(si, Tx)) {
   float theoretical = Distance(orig, other);
   float3 dir = normalize(other - orig);
-  orig += dir*(0.001f);
[…]

### 2017-12-05 raycastkernel.c

@@ -209,38 +209,8 @@ Ray Camera_Ray(Camera* camera, SceneInfo* si) {
   Ray ray;
 
   /* // ------ DOF & antialiasing ----- */
-  float3 dof_puv = (float3)(puv, fov_r);
-  float3 ray_dir = dof_puv.x*cam_right + dof_puv.y*cam_up + fov_r*cam_front;
-  float3 dof_origin = camera->radius*(float3)(Sample_Uniform(si), 0.0f, 0.0f);
-  float3 dof_dir = normalize(dof_puv * camera->focal - dof_origin);
-  cam_pos += dof_origin.x*cam_right + dof_origin.y*cam_up;
-  ray_dir += dof_dir.x*cam_right + dof_dir.y*cam_up;
-  ray_dir = normalize(ray_dir);
-  ray.origin = cam_pos;
-  ray.dir = ray_dir;
-  return ray;
-}
-
-
-typedef struct _T_EvalPreviousOutputHorcrux {
-  float4 old_colour;
-  bool raycast_nav;
-} _EvalPreviousOutputHorcrux;
-
-_EvalPreviousOutputHorcrux Eval_Previous_Output(
-                                                __global unsigned char* img, __write_only image2d_t output, int2 out,
-                                                __global SharedInfo* shared_info, int pt) {
-  _EvalPreviousOutputHorcrux hor;
-  hor.old_colour = (float4)(0.0f);
-  hor.raycast_nav = (shared_info->clear_img);
-  if ( !hor.raycast_nav) {
-    hor.old_colour = (float4)(img[pt+0]/255.0f, img[pt+1]/255.0f,
-                              img[pt+2]/255.0f, (float)(img[pt+3]));
-  } else {
-    img[pt+0] = 0.0f; img[pt+1] = 0.0f; img[pt+2] = 0.0f; img[pt+3] = 0.0f;
-    shared_info->finished_samples = 0;
-  }
-  return hor;
+  float3 ray_dir = puv.x*cam_right + puv.y*cam_up + fov_r*cam_front;
+  return (Ray){cam_pos, normalize(ray_dir)};
 }
 
 // -----------------------------------------------------------------------------
@@ -284,38 +254,14 @@ __kernel void DTOADQ_Kernel (
         float3 albedo = RColour(pt.colour, &m);
         // diffusive component i made up, no normal needed
         float3 V = -ray.dir, L = normalize(emit.origin-O);
[…]
